---
title: "NADH and NADPH metabolism"
author: "John Santiago"
date: "2025-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, message = F}

# Conditions
#Geno is WT or GR
#TKT is either "C", "OE", or "Df"
#Sex is "F" or "M"

Geno1 = "GR"
TKT1 = "C"
Sex1 = "F"

Geno2 = "WT"
TKT2 = "C"
Sex2 = "F"

##Conditions for second comparison
Geno3 = "GR"
TKT3 = "C"
Sex3 = "M"

Geno4 = "WT"
TKT4 = "C"
Sex4 = "M"





```

```{r echo = F, message = F, warning = F}


source('https://raw.githubusercontent.com/johncsantiago/WhartonLab/refs/heads/master/G85R.GOandKEGGPlottingFunctions.R')

plot.data = G85R.data(Geno1, Geno2, Geno3, Geno4,
                      Sex1, Sex2, Sex3, Sex4,
                      TKT1, TKT2, TKT3, TKT4)

```

## NADH Genes (Females)
```{r echo = F, warning = F}

##Uses data from primary comparison
##Plots genes with an FDR <= 0.05
##point size is relative to maximum mean expression in the two conditions

NADH.FBgn = row.names(read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/NADHGenes.csv", row.names = 1))

G85R.volcano(plot.data, subset = NADH.FBgn)

```


##  NADPH Genes (Females)
```{r echo = F, warning = F}

NADPH.FBgn = row.names(read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/NADPHGenes.csv", row.names = 1))

G85R.volcano(plot.data, subset = NADPH.FBgn)

```




```{r echo = F, message = F, warning = F}

plot.data = G85R.data(Geno3, Geno4, Geno1, Geno2,
                      Sex3, Sex4, Sex1, Sex2,
                      TKT3, TKT4, TKT1, TKT2)

```

## NADH Genes (Males)
```{r echo = F, warning = F}


G85R.volcano(plot.data, subset = NADH.FBgn)

```


##  NADPH Genes (Males)
```{r echo = F, warning = F}


G85R.volcano(plot.data, subset = NADPH.FBgn)

```










```{r echo = F, message = F}

# Conditions
#Geno is WT or GR
#TKT is either "C", "OE", or "Df"
#Sex is "F" or "M"

Geno1 = "GR"
TKT1 = "Df"
Sex1 = "F"

Geno2 = "GR"
TKT2 = "C"
Sex2 = "F"

##Conditions for second comparison
Geno3 = "GR"
TKT3 = "Df"
Sex3 = "M"

Geno4 = "GR"
TKT4 = "C"
Sex4 = "M"





```

```{r echo = F, message = F, warning = F}


plot.data = G85R.data(Geno1, Geno2, Geno3, Geno4,
                      Sex1, Sex2, Sex3, Sex4,
                      TKT1, TKT2, TKT3, TKT4)

```

## NADH Genes (Females)
```{r echo = F, warning = F}

G85R.volcano(plot.data, subset = NADH.FBgn)

```


##  NADPH Genes (Females)
```{r echo = F, warning = F}

G85R.volcano(plot.data, subset = NADPH.FBgn)

```




```{r echo = F, message = F, warning = F}

plot.data = G85R.data(Geno3, Geno4, Geno1, Geno2,
                      Sex3, Sex4, Sex1, Sex2,
                      TKT3, TKT4, TKT1, TKT2)

```

## NADH Genes (Males)
```{r echo = F, warning = F}


G85R.volcano(plot.data, subset = NADH.FBgn)

```


##  NADPH Genes (Males)
```{r echo = F, warning = F}


G85R.volcano(plot.data, subset = NADPH.FBgn)

```





```{r echo = F, message = F}

# Conditions
#Geno is WT or GR
#TKT is either "C", "OE", or "Df"
#Sex is "F" or "M"

Geno1 = "WT"
TKT1 = "Df"
Sex1 = "F"

Geno2 = "WT"
TKT2 = "C"
Sex2 = "F"

##Conditions for second comparison
Geno3 = "WT"
TKT3 = "Df"
Sex3 = "M"

Geno4 = "WT"
TKT4 = "C"
Sex4 = "M"





```

```{r echo = F, message = F, warning = F}


plot.data = G85R.data(Geno1, Geno2, Geno3, Geno4,
                      Sex1, Sex2, Sex3, Sex4,
                      TKT1, TKT2, TKT3, TKT4)

```

## NADH Genes (Females)
```{r echo = F, warning = F}

G85R.volcano(plot.data, subset = NADH.FBgn)

```


##  NADPH Genes (Females)
```{r echo = F, warning = F}

G85R.volcano(plot.data, subset = NADPH.FBgn)

```




```{r echo = F, message = F, warning = F}

plot.data = G85R.data(Geno3, Geno4, Geno1, Geno2,
                      Sex3, Sex4, Sex1, Sex2,
                      TKT3, TKT4, TKT1, TKT2)

```

## NADH Genes (Males)
```{r echo = F, warning = F}


G85R.volcano(plot.data, subset = NADH.FBgn)

```


##  NADPH Genes (Males)
```{r echo = F, warning = F}


G85R.volcano(plot.data, subset = NADPH.FBgn)

```









```{r echo = F, message = F}

# Conditions
#Geno is WT or GR
#TKT is either "C", "OE", or "Df"
#Sex is "F" or "M"

Geno1 = "GR"
TKT1 = "OE"
Sex1 = "F"

Geno2 = "GR"
TKT2 = "C"
Sex2 = "F"

##Conditions for second comparison
Geno3 = "GR"
TKT3 = "OE"
Sex3 = "M"

Geno4 = "GR"
TKT4 = "C"
Sex4 = "M"





```

```{r echo = F, message = F, warning = F}


plot.data = G85R.data(Geno1, Geno2, Geno3, Geno4,
                      Sex1, Sex2, Sex3, Sex4,
                      TKT1, TKT2, TKT3, TKT4)

```

## NADH Genes (Females)
```{r echo = F, warning = F}

G85R.volcano(plot.data, subset = NADH.FBgn)

```


##  NADPH Genes (Females)
```{r echo = F, warning = F}

G85R.volcano(plot.data, subset = NADPH.FBgn)

```




```{r echo = F, message = F, warning = F}

plot.data = G85R.data(Geno3, Geno4, Geno1, Geno2,
                      Sex3, Sex4, Sex1, Sex2,
                      TKT3, TKT4, TKT1, TKT2)

```

## NADH Genes (Males)
```{r echo = F, warning = F}


G85R.volcano(plot.data, subset = NADH.FBgn)

```


##  NADPH Genes (Males)
```{r echo = F, warning = F}


G85R.volcano(plot.data, subset = NADPH.FBgn)

```





```{r echo = F, message = F}

# Conditions
#Geno is WT or GR
#TKT is either "C", "OE", or "Df"
#Sex is "F" or "M"

Geno1 = "WT"
TKT1 = "OE"
Sex1 = "F"

Geno2 = "WT"
TKT2 = "C"
Sex2 = "F"

##Conditions for second comparison
Geno3 = "WT"
TKT3 = "OE"
Sex3 = "M"

Geno4 = "WT"
TKT4 = "C"
Sex4 = "M"





```

```{r echo = F, message = F, warning = F}


plot.data = G85R.data(Geno1, Geno2, Geno3, Geno4,
                      Sex1, Sex2, Sex3, Sex4,
                      TKT1, TKT2, TKT3, TKT4)

```

## NADH Genes (Females)
```{r echo = F, warning = F}

G85R.volcano(plot.data, subset = NADH.FBgn)

```


##  NADPH Genes (Females)
```{r echo = F, warning = F}

G85R.volcano(plot.data, subset = NADPH.FBgn)

```




```{r echo = F, message = F, warning = F}

plot.data = G85R.data(Geno3, Geno4, Geno1, Geno2,
                      Sex3, Sex4, Sex1, Sex2,
                      TKT3, TKT4, TKT1, TKT2)

```

## NADH Genes (Males)
```{r echo = F, warning = F}


G85R.volcano(plot.data, subset = NADH.FBgn)

```


##  NADPH Genes (Males)
```{r echo = F, warning = F}


G85R.volcano(plot.data, subset = NADPH.FBgn)

```
