---
title: "Russo Manuscript Final Figures"
author: "John Santiago"
date: "2026-06-08"
output: 
  bookdown::html_document2:
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## <span style="text-decoration:underline">Table of Contents</span> 

<a href="Russo_Manuscript_Data.html" target="_blank">Raw Data Files</a>


### Figure 1 

- #### <a href="#fig1b">Figure 1B: Model Eclosion</a>

- #### <a href="#fig1c">Figure 1C: Pharate Boutons</a>

- #### <a href="#fig1f">Figure 1F: A4V Female Survival</a>

- #### <a href="#fig1g">Figure 1G: A4V Male Survival</a>

<br /><br />

### Figure 2

- #### <a href="#fig2a">Figure 2A: A4V Female Climbing</a>

- #### <a href="#fig2b">Figure 2B: A4V Male Climbing</a>

- #### <a href="#fig2c">Figure 2C: A4V Female Axonal Branchpoints</a>

- #### <a href="#fig2d">Figure 2D: A4V Male Axonal Branchpoints</a>

- #### <a href="#fig2e">Figure 2E: A4V Female BRP Puncta</a>

- #### <a href="#fig2f">Figure 2F: A4V Male BRP Puncta</a>

- #### <a href="#fig2g">Figure 2G: Pharate Leg Morphology Scores</a>

- #### <a href="#fig2h">Figure 2H: Pharate Leg Morphology Summary</a>


<br /><br />

### Figure 3

- #### <a href="#fig3a">Figure 3A: G85R gbb and SaxA eclosion</a>

- #### <a href="#fig3b">Figure 3B: G85R gbb and SaxA Segment 4 Boutons</a>

- #### <a href="#fig3c">Figure 3C: G85R gbb and SaxA Segment 5 Boutons</a>

- #### <a href="#fig3d">Figure 3D: G85R gbb and SaxA VNC pMad Fluorescence</a>

<br /><br />

### Figure 4

- #### <a href="#fig4a">Figure 4A: A4V Female ChAT>gbb Climbing</a>

- #### <a href="#fig4b">Figure 4B: A4V Female OK371>gbb Climbing</a>

- #### <a href="#fig4c">Figure 4C: A4V Male ChAT>gbb Climbing</a>

- #### <a href="#fig4d">Figure 4D: A4V Male OK371>gbb Climbing</a>

- #### <a href="#fig4e">Figure 4E: A4V Female ChAT>SaxA Climbing</a>

- #### <a href="#fig4f">Figure 4F: A4V Female OK371>SaxA Climbing</a>

- #### <a href="#fig4g">Figure 4G: A4V Male ChAT>SaxA Climbing</a>

- #### <a href="#fig4h">Figure 4H: A4V Male OK371>SaxA Climbing</a>

<br /><br /><br />
\newpage

```{r echo = F, warning=F,message=F}

library(ggplot2)
library(gplots)
library(ggridges)
library(beeswarm)
library(kableExtra)
library(emmeans)
library(cluster)
library(vegan)
library(lme4)
library(lmerTest)
library(survival)
library(ComplexHeatmap)

std.colors = setNames(c('navy', "gold2", "turquoise4", "royalblue",
                        'orange', "deeppink", "grey", "green4"),
                      c("w1118", "WT(LoxP)", "G85R(hr)", "WT(cr)",
                        "G85R(cr)", "A4V", "A4V/X16", "A4V/G85R(cr)"))

lcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")

hex = function(color, alpha){
  rgb2hex = color
  i = 1
  while(i <= length(color)){
    rgbvals = col2rgb(color[i])
    rgb2hex[i] = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
    i = i+1
  }
  return(rgb2hex)
}

```

<a id="fig1b"></a>

## Figure 1B: G85R and A4V models eclosion data

```{r fig1b, echo = F, fig.width=10, fig.height=7}


fig1b.data = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Russo Fig 1 eclosion full data.csv")

percent.fig1b = setNames(rep(NA_real_,length(unique(fig1b.data$Genotype))),
                         unique(fig1b.data$Genotype))

#adds up all eclosed and not eclosed across replicates before calculating percent eclosed
i=1
while(i<=length(percent.fig1b)){
  tempEclosed = sum(fig1b.data[fig1b.data$Genotype == names(percent.fig1b)[i], "Eclosed"])
  tempNotEclosed = sum(fig1b.data[fig1b.data$Genotype == names(percent.fig1b)[i], "NotEclosed"])
  tempTotal = tempEclosed+tempNotEclosed
  percent.fig1b[i] = signif(100*(tempEclosed/tempTotal),3)
  i=i+1
}

#plots the total percent eclosed
par(mar = c(9,4,4,2))

barplot(height = percent.fig1b,
        names = names(percent.fig1b),
        ylim = c(0, 120),
        col = "black",
        las = 2,
        ylab = "Eclosion frequency (%)",
        main = "Eclosion from the pupal case",
        xaxt = "n",
        yaxt = "n")
axis(side = 1,
     at = c(.7 + (c(0:9)*1.2)),
     lwd = 3,
     line = 0,
     labels = names(percent.fig1b),
     xpd = T,
     las = 2)
axis(side = 2,
     at = c(0,20,40,60,80,100),
     lwd = 3,
     line = 0,
     labels = T,
     xpd = T,
     las = 2)

lines(x = c(-0.5,13.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)


lines(x = c(1.9,3.1),
      y = c(105, 105),
      lwd = 3)

lines(x = c(1.9,1.9),
      y = c(105, 102),
      lwd = 3)


text(x = 3.1,
     y = 109,
    labels = "*",
    cex = 2)


lines(x = c(4.3,11.5),
      y = c(105, 105),
      lwd = 3)

lines(x = c(4.3,4.3),
      y = c(105, 102),
      lwd = 3)


text(x = 5.5,
     y = 109,
    labels = "*",
    cex = 2)

text(x = 6.7,
     y = 109,
    labels = "*",
    cex = 2)

text(x = 9.1,
     y = 109,
    labels = "*",
    cex = 2)

text(x = 11.5,
     y = 109,
    labels = "*",
    cex = 2,
    xpd = t)

```

**Figure 1B** * indicate a Benjamini-Hochberg adjusted Fishers exact test p-value less than 0.05. <a href="https://docs.google.com/spreadsheets/d/1JB9swADR91_tpwvsZwVRsT4WoxX39LU9oX8dPJTUpAg/edit?gid=1599960449#gid=1599960449" target="_blank">Fig1B Data</a><br><br>

```{r echo = F, include = F}

#Why Fisher’s exact test is the right choice for the situation:

#Binomial outcome (eclosed vs not) 
#Comparing pairs of genotypes (mutant vs matched control) 
#Unequal sample sizes 
#Extreme values (0% / 100%) 
#Moderate number of replicates (4–9 vials) 
#Clear biological grouping (not all comparisons meaningful) 

#This is exactly the kind of scenario Fisher’s test was designed for.

##Fishers Exact Test
rawB <- aggregate(cbind(Eclosed, NotEclosed) ~ Genotype, data = fig1b.data, sum)

# total and proportion
rawB$total <- rawB$Eclosed + rawB$NotEclosed
rawB$prop  <- rawB$Eclosed / rawB$total

comparisons <- list(
  c("WTloxP", "G85Rhr"),
  c("WTcr", "G85Rcr"),
  c("WTcr", "G85R/X16"),
  c("WTcr", "WTcr/X16"),
  c("WTcr", "A4V"),
  c("WTcr", "A4V/X16"),
  c("WTcr", "A4V/G85Rcr")
)

results <- data.frame()

for (pair in comparisons) {
  
  g1 <- pair[1]
  g2 <- pair[2]
  
  d1 <- rawB[rawB$Genotype == g1, ]
  d2 <- rawB[rawB$Genotype == g2, ]
  
  mat <- matrix(c(
    d1$Eclosed, d1$NotEclosed,
    d2$Eclosed, d2$NotEclosed
  ), nrow = 2, byrow = TRUE)
  
  ft <- fisher.test(mat)
  
  results <- rbind(results, data.frame(
    Genotype1 = g1,
    n1 = d1$total,
    prop1 = d1$prop,
    Genotype2 = g2,
    n2 = d2$total,
    prop2 = d2$prop,
    diff = d1$prop - d2$prop,
    p.value = ft$p.value
  ))
}

results$padj <- p.adjust(results$p.value, method = "BH")

results$prop1 <- round(results$prop1 * 100, 1)
results$prop2 <- round(results$prop2 * 100, 1)
results$diff  <- round(results$diff * 100, 1)
results$p.value <- signif(results$p.value, 3)
results$padj    <- signif(results$padj, 3)

results

```

<a id="fig1c"></a>

## Figure 1C: Pharate bouton data

```{r fig1c, echo = F, fig.width=10, fig.height=7}

C = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1C pharate_bouton_counts.xlsx - counts.csv")

W1118 = data.frame(Genotype = "W1118",
                   Boutons = na.omit(C$W1118))


SILENT = data.frame(Genotype = "SILENT",
                   Boutons = na.omit(C$SILENT))


G85Rcr = data.frame(Genotype = "G85Rcr",
                   Boutons = na.omit(C$G85Rcr))


A4V = data.frame(Genotype = "A4V",
                   Boutons = na.omit(C$A4V))



C = rbind(W1118, rbind(SILENT,
          rbind(G85Rcr, A4V)))

C$Genotype = factor(C$Genotype, levels = c("W1118", "SILENT", "G85Rcr", "A4V"))

C.labels = c("w1118", "WT(cr)", "G85R(cr)", "A4V")
C.cols = std.colors[C.labels]
     
 
par(mar = c(5,7,5,2))
      
plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 150),
     xlim = c(.5, 4.5),
     xaxt = "n",
     ylab="Bouton Number",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Pharate muscle VM NMJ bouton counts")
      
boxplot(C$Boutons~C$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 150),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(C.cols, .25),
        add = T, 
        xaxt = "n",
        frame = F)

     
beeswarm(C$Boutons~C$Genotype,
         at = c(1,2,3,4),
       pch = 21,
       cex = 2.0,
       spacing = 1.20,
       method = "swarm",
       bg = C.cols,
       add = T)

axis(side = 1,
     at = c(1:4),
     lwd = 3,
     line = -.9,
     labels = C.labels,
     xpd = T)


axis(side = 2,
     at = c(0, 50, 100, 150),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)

lines(x = c(0.35,4.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

lines(x = c(1, 4),
     y = c(140,140),
     lwd = 3)

lines(x = c(3, 3),
     y = c(60,140),
     lwd = 3)

lines(x = c(1, 1),
     y = c(120,140),
     lwd = 3)

lines(x = c(2, 2),
     y = c(120,140),
     lwd = 3)

lines(x = c(4, 4),
     y = c(120,140),
     lwd = 3)

text(x = 3,
     y = 145,
    labels = "*",
    cex = 2)


```

**Figure 1C** * indicates a Benjamini-Hochberg adjusted p-value less than 0.05 using a pairwise Wilcoxon Rank Sum Test. <a href="https://docs.google.com/spreadsheets/d/1kte-aTJmHrUgXP-lBtj-d0_nARMdw3ORyR_tn59Gm8Q/edit?usp=sharing" target="_blank">Fig1C Data</a><br><br>

```{r echo = F, warning=F, include = F}

#Kruskal-Wallis test
#Non-parametric → no normality assumption
#Works well for small sample sizes
#Robust to unequal variance
#Standard for this kind of biological data

SummaryStatsC = aggregate(Boutons ~ Genotype, data = C, function(x) {
  c(n = length(x),
    median = median(x),
    mean = mean(x),
    sd = sd(x))
})

Summary =  data.frame(SummaryStatsC$Boutons)

row.names(Summary) = SummaryStatsC$Genotype

Summary

kruskal.test(Boutons ~ Genotype, data = C)

pairwise.wilcox.test(C$Boutons, C$Genotype, p.adjust.method = "BH")

print("There was a significant difference in the number of boutons detected in G85R with every other group tested.")
```

<a id="fig1f"></a>

## Figure 1F: A4V female survival data

```{r fig1f, echo = F, fig.width=10, fig.height=6}

fig1f = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1F left_females.csv", header = T)
fig1f.new = fig1f

w1118 = na.omit(fig1f[,2])
w1118 = c(100,w1118[1:(length(w1118)-1)])-w1118[1:length(w1118)]
w1118.deaths = w1118/min(w1118[w1118>0])
w1118 = 100/min(w1118[w1118>0])
fig1f.new[1:length(w1118.deaths),2] = w1118.deaths

dsodWT = na.omit(fig1f[,3])
dsodWT = c(100,dsodWT[1:(length(dsodWT)-1)])-dsodWT[1:length(dsodWT)]
dsodWT.deaths = dsodWT/min(dsodWT[dsodWT>0])
dsodWT = 100/min(dsodWT[dsodWT>0])
fig1f.new[1:length(dsodWT.deaths),3] = dsodWT.deaths

A4V = na.omit(fig1f[,4])
A4V = c(100,A4V[1:(length(A4V)-1)])-A4V[1:length(A4V)]
A4V.deaths = A4V/min(A4V[A4V>0])
A4V = 100/min(A4V[A4V>0])
fig1f.new[1:length(A4V.deaths),4] = A4V.deaths

A4V.X16 = na.omit(fig1f[,5])
A4V.X16 = c(100,A4V.X16[1:(length(A4V.X16)-1)])-A4V.X16[1:length(A4V.X16)]
A4V.X16.deaths = A4V.X16/min(A4V.X16[A4V.X16>0])
A4V.X16 = 100/min(A4V.X16[A4V.X16>0])
fig1f.new[1:length(A4V.X16.deaths),5] = A4V.X16.deaths

presumed.n = setNames(as.integer(c(w1118, dsodWT, A4V, A4V.X16)),
                      c("w1118", "dsodWT", "A4V", "A4V.X16"))

i=1
w1118.times = numeric()
while(i<=length(w1118.deaths)){
  temp = rep(fig1f.new$day[i], fig1f.new$w1118[i])
  w1118.times = c(w1118.times, temp)
  i=i+1
}

i=1
dsodWT.times = numeric()
while(i<=length(dsodWT.deaths)){
  temp = rep(fig1f.new$day[i], fig1f.new$dsodWT[i])
  dsodWT.times = c(dsodWT.times, temp)
  i=i+1
}

i=1
A4V.times = numeric()
while(i<=length(A4V.deaths)){
  temp = rep(fig1f.new$day[i], fig1f.new$A4V[i])
  A4V.times = c(A4V.times, temp)
  i=i+1
}

i=1
A4V.X16.times = numeric()
while(i<=length(A4V.X16.deaths)){
  temp = rep(fig1f.new$day[i], fig1f.new$A4V.X16[i])
  A4V.X16.times = c(A4V.X16.times, temp)
  i=i+1
}

notcounted = as.integer(presumed.n) - c(length(w1118.times),
                            length(dsodWT.times),
                            length(A4V.times),
                            length(A4V.X16.times))
names(notcounted) = names(presumed.n)


surv.fig1f = data.frame(Genotype = c(rep("w1118",
                                     length(w1118.times)),
                                 rep("dsodWT",
                                     length(dsodWT.times)),
                                 rep("A4V",
                                     length(A4V.times)),
                                 rep("A4V.X16",
                                     length(A4V.X16.times))),
                    Time = c(w1118.times,
                             dsodWT.times,
                             A4V.times,
                             A4V.X16.times),
                    Status = 1)

still.alive = data.frame(Genotype = rep("A4V", 3),
                         Time = rep(max(subset(surv.fig1f, Genotype == "A4V")[,"Time"]), 3),
                         Status = 0)

surv.fig1f = rbind(surv.fig1f, still.alive)

surv.fig1f$Genotype = factor(surv.fig1f$Genotype,
                                    levels = c('dsodWT',
                                               'w1118',
                                               "A4V",
                                               'A4V.X16'))

fit <- survfit(Surv(Time, Status) ~ Genotype, data = surv.fig1f)

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(fit,
     fun = "pct",
     col = c('deepskyblue', 'navy', 'deeppink2', 'grey70'),
     ylab = "Percent alive",
     ylim = c(0, 100),
     xlim = c(1,80),
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     lwd = 2,
     main = "Females",
     frame.plot = F)


axis(1, at = 20*(0:4), label = F, lwd = 3, line = -0.75)
axis(1, at = 20*(0:4), tick = F, line = -0.75)

axis(2, at = 25*(0:4), label = F, lwd = 3, line = -0.8)
axis(2, at = 25*(0:4), tick = F, las = 1, line = -0.8)

lines(x=c(0,0),
      y=c(0,100),
      lwd = 3,
      xpd = T)

lines(x=c(0,85),
      y=c(0,0),
      lwd = 3,
      xpd = T)


legend("topright",
       inset=c(-0.075,0),
       legend = c(paste0('w1118 (n = ', presumed.n["w1118"], ")"), 
                  paste0('WT(cr) (n = ', presumed.n["dsodWT"], ")"),
                  paste0('A4V (n = ', presumed.n["A4V"], ")"),
                  paste0('A4V/X16 (n = ', presumed.n["A4V.X16"], ")")),
       pch = 21,
       pt.bg = c(hex(lcol[1], .5), hex(lcol[2], .5), hex(lcol[3], .5),
                hex(lcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```

**Figure 1F** Kaplan-Meier survival curves. Used a Cox proportional hazards regression model with Holm-adjusted pairwise comparisons to assess relative hazard ratios between each genotype. A4V and A4V.X16 did not have significantly different survival from each other but did have very significantly lower survival rates than WT(cr) (WT(cr)-A4V HR = .0943, or, A4V-WT(cr) HR = 10.6 (1/.0943) meaning it is 10.6 fold more likely to die at any point. <a href="https://docs.google.com/spreadsheets/d/1QNW9TlTqpkNRRnMRU0K_ygGxSsqmo3d4yW4vjq2EKtM/edit?usp=sharing" target="_blank">Fig1F Data</a><br><br>

```{r echo = F, include = F}

cox <- coxph(Surv(Time, Status) ~ Genotype, data = surv.fig1f)
##hazard ratio is the exp(coef). For example, if w1118 has a exp(coef) of 1.36 compared to dsodWT, then that means w1118 has a 36% higher hazard of dying at any point than the dsodWT
summary(cox)

##Use cox.zph(cox) to make sure there is no significant violation of the model to run a coxph

emm <- emmeans(cox, ~ Genotype)
pw <- pairs(emm, adjust = "holm")

summary(pw, infer = c(TRUE, TRUE), type = "response")

print("Used a Cox proportional hazards regression model with Holm-adjusted pairwise comparisons to assess relative hazard ratios between each genotype. A4V and A4V.X16 did not have significantly different survival from each other but did have very significantly lower survival rates than WT(cr) (WT(cr)-A4V HR = .0693, or, A4V-WT(cr) HR = 14.43 (1/.0693) meaning it is 14.43 fold more likely to die at any point.)")

```

<a id="fig1g"></a>

## Figure 1G: A4V Male survival data

```{r fig1g, echo = F, fig.width=10, fig.height=6}

fig1g = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1G left_males.corrected.csv", header = T)
fig1g.new = fig1g

w1118 = na.omit(fig1g[,2])
w1118 = c(100,w1118[1:(length(w1118)-1)])-w1118[1:length(w1118)]
w1118.deaths = w1118/min(w1118[w1118>0])
w1118 = 100/min(w1118[w1118>0])
fig1g.new[1:length(w1118.deaths),2] = w1118.deaths

dsodWT = na.omit(fig1g[,3])
dsodWT = c(100,dsodWT[1:(length(dsodWT)-1)])-dsodWT[1:length(dsodWT)]
dsodWT.deaths = dsodWT/min(dsodWT[dsodWT>0])
dsodWT = 100/min(dsodWT[dsodWT>0])
fig1g.new[1:length(dsodWT.deaths),3] = dsodWT.deaths

A4V = na.omit(fig1g[,4])
A4V = c(100,A4V[1:(length(A4V)-1)])-A4V[1:length(A4V)]
A4V.deaths = A4V/min(A4V[A4V>0])
A4V = 100/min(A4V[A4V>0])
fig1g.new[1:length(A4V.deaths),4] = A4V.deaths

A4V.X16 = na.omit(fig1g[,5])
A4V.X16 = c(100,A4V.X16[1:(length(A4V.X16)-1)])-A4V.X16[1:length(A4V.X16)]
A4V.X16.deaths = A4V.X16/min(A4V.X16[A4V.X16>0])
A4V.X16 = 100/min(A4V.X16[A4V.X16>0])
fig1g.new[1:length(A4V.X16.deaths),5] = A4V.X16.deaths

presumed.n = setNames(as.integer(c(w1118, dsodWT, A4V, A4V.X16)),
                      c("w1118", "dsodWT", "A4V", "A4V.X16"))

i=1
w1118.times = numeric()
while(i<=length(w1118.deaths)){
  temp = rep(fig1g.new$day[i], fig1g.new$w1118[i])
  w1118.times = c(w1118.times, temp)
  i=i+1
}

i=1
dsodWT.times = numeric()
while(i<=length(dsodWT.deaths)){
  temp = rep(fig1g.new$day[i], fig1g.new$dsodWT[i])
  dsodWT.times = c(dsodWT.times, temp)
  i=i+1
}

i=1
A4V.times = numeric()
while(i<=length(A4V.deaths)){
  temp = rep(fig1g.new$day[i], fig1g.new$A4V[i])
  A4V.times = c(A4V.times, temp)
  i=i+1
}

i=1
A4V.X16.times = numeric()
while(i<=length(A4V.X16.deaths)){
  temp = rep(fig1g.new$day[i], fig1g.new$A4V.X16[i])
  A4V.X16.times = c(A4V.X16.times, temp)
  i=i+1
}

notcounted = as.integer(presumed.n) - c(length(w1118.times),
                            length(dsodWT.times),
                            length(A4V.times),
                            length(A4V.X16.times))
names(notcounted) = names(presumed.n)


surv.fig1g = data.frame(Genotype = c(rep("w1118",
                                     length(w1118.times)),
                                 rep("dsodWT",
                                     length(dsodWT.times)),
                                 rep("A4V",
                                     length(A4V.times)),
                                 rep("A4V.X16",
                                     length(A4V.X16.times))),
                    Time = c(w1118.times,
                             dsodWT.times,
                             A4V.times,
                             A4V.X16.times),
                    Status = 1)

still.alive = data.frame(Genotype = rep("A4V", 3),
                         Time = rep(max(subset(surv.fig1g, Genotype == "A4V")[,"Time"]), 3),
                         Status = 0)

surv.fig1g = rbind(surv.fig1g, still.alive)

surv.fig1g$Genotype = factor(surv.fig1g$Genotype,
                                    levels = c('dsodWT',
                                               'w1118',
                                               "A4V",
                                               'A4V.X16'))

fit <- survfit(Surv(Time, Status) ~ Genotype, data = surv.fig1g)

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(fit,
     fun = "pct",
     col = c('deepskyblue', 'navy', 'deeppink2', 'grey70'),
     ylab = "Percent alive",
     ylim = c(0, 100),
     xlim = c(1,80),
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     lwd = 2,
     main = "Males",
     frame.plot = F)


axis(1, at = 20*(0:4), label = F, lwd = 3, line = -0.75)
axis(1, at = 20*(0:4), tick = F, line = -0.75)

axis(2, at = 25*(0:4), label = F, lwd = 3, line = -0.8)
axis(2, at = 25*(0:4), tick = F, las = 1, line = -0.8)

lines(x=c(0,0),
      y=c(0,100),
      lwd = 3,
      xpd = T)

lines(x=c(0,85),
      y=c(0,0),
      lwd = 3,
      xpd = T)


legend("topright",
       inset=c(-0.075,0),
       legend = c(paste0('w1118 (n = ', presumed.n["w1118"], ")"), 
                  paste0('WT(cr) (n = ', presumed.n["dsodWT"], ")"),
                  paste0('A4V (n = ', presumed.n["A4V"], ")"),
                  paste0('A4V/X16 (n = ', presumed.n["A4V.X16"], ")")),
       pch = 21,
       pt.bg = c(hex(lcol[1], .5), hex(lcol[2], .5), hex(lcol[3], .5),
                hex(lcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```

**Figure 1G** Kaplan-Meier survival curves. Used a Cox proportional hazards regression model with Holm-adjusted pairwise comparisons to assess relative hazard ratios between each genotype. A4V and A4V.X16 did not have significantly different survival from each other but did have very significantly lower survival rates than WT(cr) (WT(cr)-A4V HR = .0194, or, A4V-WT(cr) HR = 51.54 (1/.0194) meaning it is 51.54 fold more likely to die at any point.) <a href="https://docs.google.com/spreadsheets/d/17qQ6o3-rutwdQvDqz9BTdgeBRLfKDTbagBFWnHL2qDs/edit?usp=sharing" target="_blank">Fig1G Data</a><br><br>

```{r echo = F, include = F}

cox <- coxph(Surv(Time, Status) ~ Genotype, data = surv.fig1g)
##hazard ratio is the exp(coef). For example, if w1118 has a exp(coef) of 1.36 compared to dsodWT, then that means w1118 has a 36% higher hazard of dying at any point than the dsodWT
summary(cox)

##Use cox.zph(cox) to make sure there is no significant violation of the model to run a coxph

emm <- emmeans(cox, ~ Genotype)
pw <- pairs(emm, adjust = "holm")

summary(pw, infer = c(TRUE, TRUE), type = "response")

print("Used a Cox proportional hazards regression model with Holm-adjusted pairwise comparisons to assess relative hazard ratios between each genotype. A4V and A4V.X16 did not have significantly different survival from each other but did have very significantly lower survival rates than WT(cr) (WT(cr)-A4V HR = .0194, or, A4V-WT(cr) HR = 51.54 (1/.0194) meaning it is 51.54 fold more likely to die at any point.)")

```

<a id="fig2a"></a>

## Figure 2A: A4V Female climbing data

```{r fig2a, echo = F}

fig2ab = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 2AB.csv", header = T)
fig2ab$Sample = paste(fig2ab$Genotype, paste(fig2ab$Sex, fig2ab$Day))

fig2ab = na.omit(fig2ab)

fig2ab$Genotype[fig2ab$Genotype == "dSodWT"] = "WT(cr)"

fig2ab = subset(fig2ab, Genotype != "w1118")

df_f <- subset(fig2ab, Sex == "F")
df_m <- subset(fig2ab, Sex == "M")

ggplot(df_f, aes(x = Day, y = Rate, color = Genotype)) +

  # points
  geom_point(alpha = 0.4, position = position_jitter(width = 0.2)) +

  # mean line
  stat_summary(fun = mean, geom = "line", linewidth = 1.2) +

  # mean points
  stat_summary(fun = mean, geom = "point", size = 2) +

  # CI ribbon
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(fill = Genotype),
               alpha = 0.2, color = NA) +

  scale_color_manual(values = c("WT(cr)" = "royalblue", "A4V" = "deeppink")) +
  scale_fill_manual(values = c("WT(cr)" = "royalblue", "A4V" = "deeppink")) +
  labs(title = "Female") +

  theme_classic()

```

**Figure 2A** Climbing plots show the mean at each time point, the individual data points overlayed, and the standard error of the mean as a colored ribbon. Climbing rate was modeled as a function of genotype, day, and their interaction. Model term significance was then analyzed using a two-way ANOVA. There was a significant genotype (p = 4.47e-11) and day effect on climbing rate, but the interaction of the two was not significant indicating that A4V and Silent differ in overall performance. <a href="https://docs.google.com/spreadsheets/d/1QD-RhvOwy5sE4NGbceQtLtfYl3RsR_UZKUg9MrTBNrA/edit?usp=sharing" target="_blank">Fig2AB Data</a><br><br>

```{r echo = F, include=F}

#2 way anova

model_f <- aov(Rate ~ Genotype * Day, data = df_f)
summary(model_f)


print("There was a significant effect of genotype on climbing velocity and a significant effect of age (day), indicating that genotypes differ in overall performance and that climbing declines over time. However, there was no significant genotype-by-day interaction, suggesting that the rate of decline over time does not differ between genotypes.")

model_WTF <- lm(Rate ~ Day, data = subset(df_f, Genotype == "WT(cr)"))

model_A4VF <- lm(Rate ~ Day, data = subset(df_f, Genotype == "A4V"))

summary(model_WTF)
summary(model_A4VF)

print("Both WT F and A4V F had a significant change in velocity over the time course")

```

<a id="fig2b"></a>

## Figure 2B: A4V Male climbing data


```{r fig2b, echo = F}

ggplot(df_m, aes(x = Day, y = Rate, color = Genotype)) +

  # points
  geom_point(alpha = 0.4, position = position_jitter(width = 0.2)) +

  # mean line
  stat_summary(fun = mean, geom = "line", linewidth = 1.2) +

  # mean points
  stat_summary(fun = mean, geom = "point", size = 2) +

  # CI ribbon
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(fill = Genotype),
               alpha = 0.2, color = NA) +

  scale_color_manual(values = c("WT(cr)" = "royalblue", "A4V" = "deeppink")) +
  scale_fill_manual(values = c("WT(cr)" = "royalblue", "A4V" = "deeppink")) +
  labs(title = "Male") +

  theme_classic()

```

**Figure 2B** Climbing plots show the mean at each time point, the individual data points overlayed, and the standard error of the mean as a colored ribbon. Climbing rate was modeled as a function of genotype, day, and their interaction. Model term significance was then analyzed using a two-way ANOVA. There was a significant genotype (p = 2.2e-16), day and interaction effect on climbing rate. Post hoc pairwise analysis of the significant interaction was done using estimated marginal means (emmeans) and comparison of the genotype specific slopes (emtrends). The estimated marginal means and genotype specific slopes were found to be significantly different between WT(cr) and A4V (p < 0.0001) indicating they significantly differ in overall performance.<br><br>

```{r echo = F, include = F}

#2 way anova
fit <- lm(Rate ~ Genotype * Day, data = df_m)

anova(fit)

print("There was a significant effect of genotype on climbing velocity, a significant effect of age (day), indicating that genotypes differ in overall performance and that climbing declines over time, and a significant genotype-by-day interaction, suggesting that the rate of decline over time differ between genotypes.")

tr = emtrends(fit, ~Genotype, var = "Day")
summary(tr)

pairs(tr, adjust = "holm")

emm <- emmeans(fit, ~ Genotype)
pairs(emm)

print("WT males did not have a significant change in velocity over the time course but A4V M did p = 9.78e-12")
```

<a id="fig2c"></a>

## Figure 2C: A4V Female Axonal Branchpoints

```{r fig2c, echo = F, fig.height=5, fig.width=10}

f2.bp = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/RussoManuscript_A4V_AxonalBranchpointData.csv")
f2.bp$Genotype = "WT"
f2.bp$Genotype[grep("A4V", f2.bp$Name)] = "A4V"

ff2.bp = subset(f2.bp, Sex == "F")
ff2.bp$Age = factor(ff2.bp$Age)
ff2.bp$Name = factor(ff2.bp$Name, levels = c("WTF3", "A4VF3",
                                             "WTF10", "A4VF10",
                                             "WTF21", "A4VF21"))

par(mar = c(5,7,5,2))
      
plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 30),
     xlim = c(.5, 8.5),
     xaxt = "n",
     ylab="Branchpoint Number",
     yaxt = "n",
     xlab = "Age",
     frame = F,
     main = "Female")
      
boxplot(ff2.bp$Branchpoints~ff2.bp$Name,
        at = c(1,2, 4,5, 7,8),
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 150),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(c("royalblue", "firebrick"), .25),
        add = T, 
        xaxt = "n",
        frame = F)

     
beeswarm(ff2.bp$Branchpoints~ff2.bp$Name,
         at = c(1,2, 4,5, 7,8),
       pch = 21,
       cex = 1.15,
       spacing = 1.2,
       method = "swarm",
       bg = c("royalblue","firebrick"),
       add = T)


axis(side = 1,
     at = c(1,4,7)+.5,
     lwd = 0,
     line = -.9,
     labels = levels(ff2.bp$Age),
     xpd = T)


axis(side = 2,
     at = c(0, 10, 20, 30),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)

lines(x = c(0.2,8.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

lines(x = c(4, 5),
     y = c(21,21),
     lwd = 2)

lines(x = c(7, 8),
     y = c(21,21),
     lwd = 2)

text(x = 7.5,
     y = 22.5,
    labels = "*",
    cex = 1.5)

text(x = 4.5,
     y = 22.5,
    labels = "*",
    cex = 1.5)

abline(v = c(3,6),
       lty = 2)

legend('topright',
       inset=c(-0.0,0),
       legend = c('WT', 'A4V'), 
       fill = c("royalblue", "firebrick"),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

```

**Figure 2C** Total axonal branchpoints were modeled as a function of genotype, age, and their interaction. Model term significance was then analyzed using a two-way ANOVA. There was a significant genotype (p = 2.248e-06), age and interaction effect on branchpoint number. Post hoc pairwise analysis of the significant genotype effect at each age was done by comparing the estimated marginal means (emmeans) and found a significant difference in total branchpoits between WT(cr) and A4V at 10 (p = 0.0013) and 21 (p = 0.0002) days old. <a href="https://docs.google.com/spreadsheets/d/1y5ohw4YEdhrokaJvHr285JhXjRiLD_pzh7Ri54J0lJ4/edit?usp=sharing" target="_blank">Fig2CD Data</a><br><br>

<a id="fig2d"></a>

## Figure 2D: A4V Male Axonal Branchpoints

```{r echo = F, include=F}

##Do branchpoints differ between genotypes at different ages. NOT using age as a linear relationship
fit <- lm(Branchpoints ~ Genotype * Age, data = ff2.bp)

anova(fit)

##compares genotype at each age instead of across all ages together like emmmeans(fit, pairwise ~ Genotype) would
emm = emmeans(fit, ~ Genotype | Age)
pairs(emm, adjust = "holm")

emm = emmeans(fit, ~ Age | Genotype)
pairs(emm, adjust = "holm")

```


```{r echo = F, fig.height=5, fig.width=7, fig2d}

mf2.bp = subset(f2.bp, Sex == "M")
mf2.bp$Age = factor(mf2.bp$Age)
mf2.bp$Name = factor(mf2.bp$Name, levels = c("WTM3", "A4VM3",
                                             "WTM10", "A4VM10"))

par(mar = c(5,7,5,2))
      
plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 30),
     xlim = c(.5, 5.5),
     xaxt = "n",
     ylab="Branchpoint Number",
     yaxt = "n",
     xlab = "Age",
     frame = F,
     main = "Male")
      
boxplot(mf2.bp$Branchpoints~mf2.bp$Name,
        at = c(1,2, 4,5),
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 30),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(c("royalblue", "firebrick"), .25),
        add = T, 
        xaxt = "n",
        frame = F)

     
beeswarm(mf2.bp$Branchpoints~mf2.bp$Name,
         at = c(1,2, 4,5),
       pch = 21,
       cex = 1.15,
       spacing = 1.2,
       method = "swarm",
       bg = c("royalblue","firebrick"),
       add = T)


axis(side = 1,
     at = c(1.5,4.5),
     lwd = 0,
     line = -.9,
     labels = levels(mf2.bp$Age),
     xpd = T)


axis(side = 2,
     at = c(0, 10, 20, 30),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)

lines(x = c(0.2,5.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

lines(x = c(4, 5),
     y = c(26,26),
     lwd = 2)


text(x = 4.5,
     y = 27.5,
    labels = "*",
    cex = 1.5)

abline(v = c(3,6),
       lty = 2)

legend('topright',
       inset=c(-0.075,0),
       legend = c('WT', 'A4V'), 
       fill = c("royalblue", "firebrick"),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

```


```{r echo = F, include = F}

##Do branchpoints differ between genotypes at different ages. NOT using age as a linear relationship
fit <- lm(Branchpoints ~ Genotype * Age, data = mf2.bp)

anova(fit)

##compares genotype at each age instead of across all ages together like emmmeans(fit, pairwise ~ Genotype) would
emm = emmeans(fit, ~ Genotype | Age)
pairs(emm, adjust = "holm")

##compares branchpoints between each age within a genotype
#emm = emmeans(fit, ~ Age | Genotype)
#pairs(emm, adjust = "holm")

```

**Figure 2D** Total axonal branchpoints were modeled as a function of genotype, age, and their interaction. Model term significance was then analyzed using a two-way ANOVA. There was a significant genotype (p = 0.001550), age and interaction effect on branchpoint number. Post hoc pairwise analysis of the significant genotype effect at each age was done by comparing the estimated marginal means (emmeans) and found a significant difference in total branchpoits between WT(cr) and A4V at 10 (p < 0.0001) days old.<br><br>

<a id="fig2e"></a>

## Figure 2E: A4V Female BRP Puncta

```{r echo = F, fig.height=5, fig.width=10, fig2e}

f2.punc = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/RussoManuscript_A4V_BRPPunctaData.csv")
f2.punc$Genotype = "WT"
f2.punc$Genotype[grep("A4V", f2.punc$Name)] = "A4V"

ff2.punc = subset(f2.punc, Sex == "F")
ff2.punc$Age = factor(ff2.punc$Age)
ff2.punc$Name = factor(ff2.punc$Name, levels = c("WTF10", "A4VF10",
                                                 "WTF21", "A4VF21"))

par(mar = c(5,7,5,2))
      
plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 250),
     xlim = c(.5, 5.5),
     xaxt = "n",
     ylab="Puncta Number",
     yaxt = "n",
     xlab = "Age",
     frame = F,
     main = "Female")
      
boxplot(ff2.punc$brp.puncta~ff2.punc$Name,
        at = c(1,2, 4,5),
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 250),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(c("royalblue", "firebrick"), .25),
        add = T, 
        xaxt = "n",
        frame = F)

     
beeswarm(ff2.punc$brp.puncta~ff2.punc$Name,
         at = c(1,2, 4,5),
       pch = 21,
       cex = 1.15,
       spacing = 1.2,
       method = "swarm",
       bg = c("royalblue","firebrick"),
       add = T)


axis(side = 1,
     at = c(1.5,4.5),
     lwd = 0,
     line = -.9,
     labels = levels(ff2.punc$Age),
     xpd = T)


axis(side = 2,
     at = c(0, 50, 100, 150, 200, 250),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)

lines(x = c(0.2,5.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

abline(v = c(3,6),
       lty = 2)

legend('topright',
       inset=c(-0.0,0),
       legend = c('WT', 'A4V'), 
       fill = c("royalblue", "firebrick"),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

```

**Figure 2E** Total bruchpilot puncta were modeled as a function of genotype, age, and their interaction. Model term significance was then analyzed using a two-way ANOVA. There was no significant genotype effect on total bruchpilot puncta. Post hoc pairwise analysis at each age was done by comparing the estimated marginal means (emmeans) and confirmed that there was not a significant difference in total bruchpilot puncta between WT(cr) and A4V at either 10 or 21 days old. <a href="https://docs.google.com/spreadsheets/d/1uPBgzL_LV95Eb_bmb-2h5AVAjCTT5MhbHhCz9d8ckZw/edit?usp=sharing" target="_blank">Fig2EF Data</a><br><br>

```{r echo = F, include = F}

##Do BRP puncta differ between genotypes at different ages. NOT using age as a linear relationship
fit <- lm(brp.puncta ~ Genotype * Age, data = ff2.punc)

anova(fit)

##compares genotype at each age instead of across all ages together like emmmeans(fit, pairwise ~ Genotype) would
emm = emmeans(fit, ~ Genotype | Age)
pairs(emm, adjust = "holm")

emm = emmeans(fit, ~ Age | Genotype)
pairs(emm, adjust = "holm")

```

<a id="fig2f"></a>

## Figure 2F: A4V Male BRP Puncta

```{r echo = F, fig.height=5, fig.width=10, fig2f}
f2.punc = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/RussoManuscript_A4V_BRPPunctaData.csv")
f2.punc$Genotype = "WT"
f2.punc$Genotype[grep("A4V", f2.punc$Name)] = "A4V"

mf2.punc = subset(f2.punc, Sex == "M")
mf2.punc$Age = factor(mf2.punc$Age)
mf2.punc$Name = factor(mf2.punc$Name, levels = c("WTM3", "A4VM3",
                                                 "WTM10", "A4VM10"))

par(mar = c(5,7,5,2))
      
plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 250),
     xlim = c(.5, 5.5),
     xaxt = "n",
     ylab="Puncta Number",
     yaxt = "n",
     xlab = "Age",
     frame = F,
     main = "Female")
      
boxplot(mf2.punc$brp.puncta~mf2.punc$Name,
        at = c(1,2, 4,5),
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 250),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(c("royalblue", "firebrick"), .25),
        add = T, 
        xaxt = "n",
        frame = F)

     
beeswarm(mf2.punc$brp.puncta~mf2.punc$Name,
         at = c(1,2, 4,5),
       pch = 21,
       cex = 1.15,
       spacing = 1.2,
       method = "swarm",
       bg = c("royalblue","firebrick"),
       add = T)


axis(side = 1,
     at = c(1.5,4.5),
     lwd = 0,
     line = -.9,
     labels = levels(mf2.punc$Age),
     xpd = T)


axis(side = 2,
     at = c(0, 50, 100, 150, 200, 250),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)

lines(x = c(0.2,5.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

lines(x = c(4, 5),
     y = c(200,200),
     lwd = 2)

text(x = 4.5,
     y = 215,
    labels = "*",
    cex = 1.5)

abline(v = c(3,6),
       lty = 2)

legend('topright',
       inset=c(-0.0,0),
       legend = c('WT', 'A4V'), 
       fill = c("royalblue", "firebrick"),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

```

```{r echo = F, include = F}

##Do BRP puncta differ between genotypes at different ages. NOT using age as a linear relationship
fit <- lm(brp.puncta ~ Genotype * Age, data = mf2.punc)

anova(fit)

##compares genotype at each age instead of across all ages together like emmmeans(fit, pairwise ~ Genotype) would
emm = emmeans(fit, ~ Genotype | Age)
pairs(emm, adjust = "holm")

emm = emmeans(fit, ~ Age | Genotype)
pairs(emm, adjust = "holm")

```

**Figure 2F** Total bruchpilot puncta were modeled as a function of genotype, age, and their interaction. Model term significance was then analyzed using a two-way ANOVA. There was a significant genotype (p = 0.002568) effect on total bruchpilot puncta. Post hoc pairwise analysis at each age was done by comparing the estimated marginal means (emmeans) and found that there was a significant difference in total bruchpilot puncta between WT(cr) and A4V at 10 (p = 0.0037) days old, but not at 3.<br><br>

<a id="fig2g"></a>

## Figure 2G: Pharate Leg Morphology Scores

```{r echo = F, warning=F, message=F, fig.width=10, fig.height=10, fig2g}

UB = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/UBLegMorphologyData.csv")

temp.titles = data.frame(Feature = colnames(UB)[6:14],
                         Title = c("Is the leg short?",
                "Is there scrunching or curving?",
                "Is the axonal bundle straight?",
                "Does the axonal bundle split or are there chunks missing?",
                "Does the branching look organized?",
                "Is the branching following a pattern tiling?",
                "Is there branching in all aspects of the femur?",
                "Is the mitochonria in the entire axonal bundle?",
                "Do all branches have mitochondria?"))

row.names(temp.titles) = temp.titles$Feature


score_cols = colnames(UB)[c(8:14)]
#print("used all categores related to the neuronal phenotypes, not those for leg morphology")

UB$score <- rowSums(UB[, score_cols], na.rm = TRUE)

#print("wilcoxon rank sum test to see if the overall scores for legs in one group are higher than those of the other group")
#wilcox.test(score ~ Genotype, data = UB)


UB$Genotype[UB$Genotype == "WT"] = "WT(cr)"
UB$Genotype[UB$Genotype == "G85R"] = "G85R(cr)"

UB$Genotype = factor(UB$Genotype, levels = c("WT(cr)", "G85R(cr)"))

morph.data = t(UB[,c("Genotype",score_cols)])


i=1
morph.plotdata = data.frame(Feature = temp.titles[temp.titles$Feature %in% row.names(morph.data)[2:nrow(morph.data)], "Feature"],
                            Genotype = rep(as.character(morph.data["Genotype",i]),
      nrow(morph.data)-1),
                            Score = morph.data[2:nrow(morph.data),i])

i=2
while(i <= ncol(morph.data)){
  temp = data.frame(Feature = temp.titles[temp.titles$Feature %in% row.names(morph.data)[2:nrow(morph.data)], "Feature"],
                    Genotype = rep(as.character(morph.data["Genotype",i]),
      nrow(morph.data)-1),
      Score = morph.data[2:nrow(morph.data),i])
  morph.plotdata = rbind(morph.plotdata, temp)
  i=i+1
}


all.morphplot = morph.plotdata
all.morphplot$Genotype = factor(all.morphplot$Genotype, levels = c("WT(cr)","G85R(cr)"))
all.morphplot$Score = as.numeric(all.morphplot$Score)
all.morphplot$Feature = factor(all.morphplot$Feature, c(all.morphplot$Feature[7:1]))

feature_results <- data.frame()

for (feature in levels(all.morphplot$Feature)) {
  
  subset.data = subset(all.morphplot, Feature == feature)
  
  tab <- table(subset.data$Genotype, subset.data$Score)
  
  ft <- fisher.test(tab)
  
  feature_results <- rbind(feature_results, data.frame(
    Feature = feature,
    p.value = ft$p.value
  ))
}

all.morphplot$Title = temp.titles[all.morphplot$Feature, "Title"]


ggplot(
  all.morphplot, 
  aes(x = `Score`, y = `Title`)
  ) +
  geom_density_ridges_gradient(
    aes(fill = Genotype), scale = 1
    ) +
  scale_fill_manual(
    values = alpha(c("royalblue", "orange"), .5)
  ) +
  labs(title = 'Morphology Score') +
  annotate("text", 
           x = rep(1.7, 7), 
           y = as.numeric(row.names(feature_results))+.8,
           label = paste0("p = ",signif(feature_results$p.value, 2)),
           cex = 2)






```

**Figure 2G** Each category related to a neuronal phenotype was individually compared using a Fishers exact test. <a href="https://docs.google.com/spreadsheets/d/1uipyv2YQoII3xojnKMDzglnXtNHg9f0cXdyR1jHwMEU/edit?usp=sharing" target="_blank">Fig2GH Data</a><br><br>

<a id="fig2h"></a>

## Figure 2H: Pharate Leg Morphology Summary

```{r echo = F, warning=F, fig.width=8, fig.height=8, fig2h}

temp = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/UBLegMorphologyData.csv")

temp$Genotype[temp$Genotype == "WT"] = "WT(cr)"
temp$Genotype[temp$Genotype == "G85R"] = "G85R(cr)"

temp$Genotype = factor(temp$Genotype, levels = unique(temp$Genotype))

morph.data = temp[,c(1:2,7:9,12:14)]
morph.data = temp[,c(1:2,8:14)]


sum.morph = setNames(apply(morph.data[,3:ncol(morph.data)], MARGIN = 1, sum), morph.data$Genotype)

#Wilcoxon rank sum test, are the values in one group higher than the other?
wrst = wilcox.test(sum.morph ~ temp$Genotype)
pval = wrst[[3]]

par(mar = c(5,7,5,2))
      
plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 7.0),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= "Morphology Score",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Leg Axon Morphology")

grid(nx=NA, ny=NULL, col = "grey50", lwd = .75)
      
boxplot(sum.morph~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 6),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        yaxt = "n",
        frame = F)


  axis(side = 1,
       at = c(1,2),
       lwd = 0,
       line = -.45,
       labels = unique(temp$Genotype),
       xpd = T)
  
    lines(x = c(0.4, 2.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)


beeswarm(sum.morph~as.numeric(temp$Genotype),
       pch = 21,
       cex = 1.5,
       spacing = 1.0,
       method = "swarm",
       bg = std.colors[as.character(unique(temp$Genotype))],
       add = T)

      axis(side = 2,
       at = c(0:7),
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)
  
  text(x = 2.25,
       y = 6.75,
       label = paste0("pval = ", signif(pval, 3)))
  
```

**Figure 2H** A morphology score was generated for each sample by summing all phenotype scores. Morphology scores were were found to be significantly different between WT and A4V (p = 0.0023) when compared using a Wilcoxon Rank Sum test.<br><br>

<a id="fig3a"></a>

## Figure 3A: G85R gbb and SaxA eclosion

```{r echo = F, fig.width=10, fig.height=7, fig3a}

gbbEcl = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Ana_gbbEclosionData.csv")

gbbEcl.labels = unique(gbbEcl$Genotype)
gbbEcl.labels = c("WT(LoxP/cr)", "OK371-gbb; WT(LoxP/cr)", "OK371-SaxA; WT(LoxP/cr)",
                  "G85R(hr/cr)", "OK371-gbb; G85R(hr/cr)", "OK371-SaxA; G85R(hr/cr)")


gbbEcl.means = rep(0, length(gbbEcl.labels))

i = 1
while(i<=length(gbbEcl.labels)){
  gbbEcl.means[i] = mean(gbbEcl$PercentEclosion[gbbEcl$Genotype == gbbEcl.labels[i]])
  i = i+1
}

stderr.gbbEcl = setNames(c(sd(gbbEcl[gbbEcl==gbbEcl.labels[1],"PercentEclosion"])/2,
                           sd(gbbEcl[gbbEcl==gbbEcl.labels[2],"PercentEclosion"])/2,
                           sd(gbbEcl[gbbEcl==gbbEcl.labels[3],"PercentEclosion"])/2,
                           sd(gbbEcl[gbbEcl==gbbEcl.labels[4],"PercentEclosion"])/2,
                           sd(gbbEcl[gbbEcl==gbbEcl.labels[5],"PercentEclosion"])/2,
                           sd(gbbEcl[gbbEcl==gbbEcl.labels[6],"PercentEclosion"])/2),
                         gbbEcl.labels)





par(mar = c(12,4,4,2))

barplot(height = gbbEcl.means,
        ylim = c(0, 120),
        col = "black",
        las = 2,
        ylab = "Eclosion frequency (%)",
        main = "Eclosion from the pupal case",
        xaxt = "n",
        yaxt = "n")
axis(side = 1,
     at = c(.7 + (c(0:5)*1.2)),
     lwd = 3,
     line = 0,
     labels = gbbEcl.labels,
     xpd = T,
     las = 2)
axis(side = 2,
     at = c(0,20,40,60,80,100),
     lwd = 3,
     line = 0,
     labels = T,
     xpd = T,
     las = 2)


tick.x = c(.7 + (c(0:5)*1.2))

i = 1
while(i <= length(stderr.gbbEcl)){
  if(stderr.gbbEcl[i] > 0){
    lines(x = tick.x[c(i,i)],
          y = c(gbbEcl.means[i], (gbbEcl.means[i] + stderr.gbbEcl[i])),
          lwd = 3,
          col = "black")
    
    lines(x = c((tick.x[i] - .25), (tick.x[i] + .25)),
          y = c((gbbEcl.means[i] + stderr.gbbEcl[i]), (gbbEcl.means[i] + stderr.gbbEcl[i])),
          lwd = 3,
          col = "black")  
  }
  i = i +1
}

lines(x = c(4.3,6.7),
      y = c(80, 80),
      lwd = 3)

lines(x = c(4.3,4.3),
      y = c(80, 77),
      lwd = 3)

text(x = c(5.5, 6.7), 
     y = c(85, 85),
     label = c("*", "*"),
     cex = 2)

```

**Figure 3A** * indicate a Benjamini-Hochberg adjusted Fishers exact test p-value less than 0.05. <a href="https://docs.google.com/spreadsheets/d/1FrXHU0W8I3hn_Ef569aX4ufS-Fl3i2y4YxgtY_Bf0Vc/edit?usp=sharing" target="_blank">Fig3A Data</a><br><br>

```{r echo = F, include = F, warning=F}

#Why Fisher’s test fits this dataset

#counts (male + female = total eclosed)
#effectively success vs failure (eclosed vs not eclosed — even if “not eclosed” isn’t explicitly listed, you can infer totals)
#small number of replicates (4 broods)
#extreme values (lots of zeros)

#This is exactly where Fisher’s test shines:

#robust to small sample sizes
#handles extreme proportions (0% vs high %)
#no distribution assumptions

##Another Fishers Test


fet.gbbEcl = gbbEcl
fet.gbbEcl$eclosed = fet.gbbEcl$TotalEclosed
fet.gbbEcl$not.eclosed = fet.gbbEcl$TotalPupae - fet.gbbEcl$TotalEclosed

fet.gbbEcl <- aggregate(cbind(eclosed, not.eclosed) ~ Genotype, data = fet.gbbEcl, sum)

# total and proportion
fet.gbbEcl$total <- fet.gbbEcl$eclosed + fet.gbbEcl$not.eclosed
fet.gbbEcl$prop  <- fet.gbbEcl$eclosed / fet.gbbEcl$total

comparisons <- list(
  c("WT(LoxP/cr)", "G85R(hr/cr)"),
  c("WT(LoxP/cr)", "OK371-gbb; WT(LoxP/cr)"),
  c("WT(LoxP/cr)", "OK371-SaxA; WT(LoxP/cr)"),
  c("G85R(hr/cr)", "OK371-gbb; G85R(hr/cr)"),
  c("G85R(hr/cr)", "OK371-SaxA; G85R(hr/cr)")
)

results <- data.frame()


#pair is like i, it starts at 1 and increases until it is larger than the length of comparisons
for (pair in comparisons) {
  
  g1 <- pair[1]
  g2 <- pair[2]
  
  d1 <- fet.gbbEcl[fet.gbbEcl$Genotype == g1, ]
  d2 <- fet.gbbEcl[fet.gbbEcl$Genotype == g2, ]
  
  mat <- matrix(c(
    d1$eclosed, d1$not.eclosed,
    d2$eclosed, d2$not.eclosed
  ), nrow = 2, byrow = TRUE)
  
  ft <- fisher.test(mat)
  
  results <- rbind(results, data.frame(
    Genotype1 = g1,
    n1 = d1$total,
    prop1 = d1$prop,
    Genotype2 = g2,
    n2 = d2$total,
    prop2 = d2$prop,
    diff = d1$prop - d2$prop,
    p.value = ft$p.value
  ))
}

results$padj <- p.adjust(results$p.value, method = "BH")

results$prop1 <- round(results$prop1 * 100, 1)
results$prop2 <- round(results$prop2 * 100, 1)
results$diff  <- round(results$diff * 100, 1)
results$p.value <- signif(results$p.value, 3)
results$padj    <- signif(results$padj, 3)

results.gbbeclosion = results

results.gbbeclosion[, c("Genotype1", "Genotype2","padj")]

```

<a id="fig3b"></a>

## Figure 3B: G85R gbb and SaxA Segment 4 Boutons

```{r echo = F, fig.width=8, fig.height=6, fig3b}
##Barron's Boutons

bb = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/20260603_BarronBoutonCounts.csv")

bb$Genotype[bb$Genotype == "WT"] = "WTcr/LoxP"
bb$Genotype[bb$Genotype == "G85R"] = "G85Rcr/hr"

bb$Genotype[bb$Genotype == "OK371>gbb WT"] = "OK371>gbb, WT"
bb$Genotype[bb$Genotype == "OK371>gbb G85R"] = "OK371>gbb, G85R"

bb$Genotype[bb$Genotype == "OK371>SaxA WT"] = "OK371>SaxA, WT"
bb$Genotype[bb$Genotype == "OK371>SaxA G85R"] = "OK371>SaxA, G85R"


bb4 = bb[bb$Segment == "A4", c('Genotype', "Boutons")]

bbplot = bb4

bbplot$color = "grey"
bbplot$color[grep("SaxA", bbplot$Genotype)] = "firebrick"
bbplot$color[grep("gbb", bbplot$Genotype)] = "steelblue"

bbplot$Genotype = factor(bbplot$Genotype, levels = c("WTcr/LoxP",
                                                     "OK371>gbb, WT",
                                                     "OK371>SaxA, WT",
                                                     "G85Rcr/hr",
                                                     "OK371>gbb, G85R",
                                                     "OK371>SaxA, G85R"))

par(mar=c(5.1, 4.1, 4.1, 6.3))

boxplot(bbplot$Boutons~bbplot$Genotype,
        at = c(1,2,3, 5,6,7),
        xaxt = 'none',
        las = 2,
        xlab = "",
        col = hex(c("grey", "steelblue", "firebrick"), .25),
        ylab = 'Total Boutons',
        main = "Segment 4 Boutons",
        ylim = c(0,60))


beeswarm(bbplot$Boutons~bbplot$Genotype,
         at = c(1,2,3, 5,6,7),
       pch = 21,
       cex = 1.0,
       spacing = 1.0,
       method = "swarm",
       bg = c("grey", "steelblue", "firebrick"),
       add = T)


axis(side = 1,
     at = c(2, 6),
     labels = c("WT(cr/hr)", "G85R(cr/hr)"))

legend('topright',
       inset=c(-0.225,0),
       legend = c('Control', 'gbb', "SaxA"), 
       fill = c("grey", "steelblue", "firebrick"),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

lines(x=c(5,7),
      y=c(60,60))

lines(x=c(6,7),
      y=c(57,57))

text(x = 6,
     y = 61,
     labels = "*")

text(x = 6.5,
     y = 58,
     labels = "*")

abline(v = 4, lty = 2)

```

**Figure 3B** Total boutons were modeled as a function of genotype, rescue, and their interaction. The model did not have a significant main effect alone but the interaction (p = 0.02) and rescue (p = 0.033) effects were significant. Post hoc pairwise analysis at each genotype*rescue was done by comparing the estimated marginal means (emmeans) and found that there was a significant difference in total boutons between the comparisons OK371>SaxA, G85R - G85R(cr/hr) (p = 0.0197) and OK371>SaxA, G85R - OK371>gbb, G85R (p = 0.0064). <a href="https://docs.google.com/spreadsheets/d/1cAdgzSjpHCE_6b-J_C4uhnlQwXV0XRiD0aKd4NG1qn8/edit?usp=sharing" target="_blank">Fig3BC Data</a><br><br>

```{r echo = F, warning=F, include=F}

print("Generated a linear model using all of the data collected on segment A4 to detect main effects of the Rescue and Model")
A4 <- subset(bb, Segment=="A4")

fit <- lm(Boutons ~ Model * Rescue, data=A4)

anova(fit)

print("The model did not have a significant main effect alone but the rescue did and the interaction effect of the model and rescue was also significant. This indicates that gbb and SaxA are significantly altering bouton number and this effect is different between the genotypes.")

emmeans(fit, pairwise ~ Model * Rescue,
        adjust="tukey")

print("Individual comparisons found that the effects of SaxA in G85R significantly altered bouton counts from G85R in the control and gbb backgrounds")

```

<a id="fig3c"></a>

## Figure 3C: G85R gbb and SaxA Segment 5 Boutons

```{r echo = F, fig.width=8, fig.height=6, fig3c}
##Barron's Boutons

bb5 = bb[bb$Segment == "A5", c('Genotype', "Boutons")]


bbplot = bb5

bbplot$color = "grey"
bbplot$color[grep("SaxA", bbplot$Genotype)] = "firebrick"
bbplot$color[grep("gbb", bbplot$Genotype)] = "steelblue"

bbplot$Genotype = factor(bbplot$Genotype, levels = c("WTcr/LoxP",
                                                     "OK371>gbb, WT",
                                                     "OK371>SaxA, WT",
                                                     "G85Rcr/hr",
                                                     "OK371>gbb, G85R",
                                                     "OK371>SaxA, G85R"))

par(mar=c(5.1, 4.1, 4.1, 6.3))

boxplot(bbplot$Boutons~bbplot$Genotype,
        at = c(1,2,3, 5,6,7),
        xaxt = 'none',
        las = 2,
        xlab = "",
        col = hex(c("grey", "steelblue", "firebrick"), .25),
        ylab = 'Total Boutons',
        main = "Segment 5 Boutons",
        ylim = c(0,60))


beeswarm(bbplot$Boutons~bbplot$Genotype,
         at = c(1,2,3, 5,6,7),
       pch = 21,
       cex = 1.0,
       spacing = 1.0,
       method = "swarm",
       bg = c("grey", "steelblue", "firebrick"),
       add = T)


axis(side = 1,
     at = c(2, 6),
     labels = c("WT(cr/hr)", "G85R(cr/hr)"))

legend('topright',
       inset=c(-0.225,0),
       legend = c('Control', 'gbb', "SaxA"), 
       fill = c("grey", "steelblue", "firebrick"),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

lines(x=c(1,5),
      y=c(60,60))

lines(x=c(1,2),
      y=c(57,57))

lines(x=c(5,7),
      y=c(57,57))

text(x = 2.5,
     y = 61,
     labels = "*")

text(x = 1.5,
     y = 58,
     labels = "*")

text(x = 6,
     y = 58,
     labels = "*")

abline(v = 4, lty = 2)

```

```{r echo = F, warning=F, include=F}

print("Generated a linear model using all of the data collected on segment A4 to detect main effects of the Rescue and Model")
A5 <- subset(bb, Segment=="A5")

fit <- lm(Boutons ~ Model * Rescue, data=A5)

anova(fit)

print("Neither the model or rescue had a significant main effect alone but the interaction effect of the model and rescue together was significant. This indicates that gbb and SaxA are significantly altering bouton number differently between the genotypes.")

emmeans(fit, pairwise ~ Model * Rescue,
        adjust="tukey")

print("Individual comparisons found that only WT and G85R have significantly different numbers of boutons at an adj. p of .05. Using a threshold of .1, the comparisons G85R ~ G85R-SaxA and WT ~ WT-gbb were found to have  significantly altered bouton counts")

```

**Figure 3C** Total boutons were modeled as a function of genotype, rescue, and their interaction. The model and rescue did not have a significant main effect alone but the interaction effect (p = 0.001293) of the model was significant. Post hoc pairwise analysis at each genotype*rescue was done by comparing the estimated marginal means (emmeans) and found that only the WT - G85R comparison has a significantly different numbers of boutons (p = 0.0376) at an adj. p of .05. Using a threshold of .1, the comparisons OK371>SaxA, G85R - G85R(cr/hr) (p = 0.0796) and OK371>gbb, WT - WT(cr/hr) (p = 0.0599) were also considered to have a significant difference.<br><br>

<a id="fig3d"></a>

## Figure 3D: G85R gbb and SaxA VNC pMad Fluorescence

```{r echo = F, fig.width=8, fig.height=6, fig3d}

temp = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/20250809_AllAnaData.csv", row.names = 1)
temp = temp[temp$Boutons != "x",]
temp$Boutons = as.numeric(temp$Boutons)
temp$Genotype[temp$Genotype == "WT-cr/LoxP"] = "WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "G85R"] = "G85R(cr/hr)"
temp$Genotype[temp$Genotype == "G85Rcr/G85Rhr"] = "G85R(cr/hr)"

temp$Genotype[temp$Genotype == "OK371 > gbb [WT]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371 > gbb [G85R]"] = "OK371-gbb; G85R(cr/hr)"
temp$Genotype[temp$Genotype == "OK371>gbb [WT]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [G85R]"] = "OK371-gbb; G85R(cr/hr)"
temp$Genotype[temp$Genotype == "OK371>gbb [WTcr/LoxP]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [WT-cr/LoxP]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [G85Rcr/G85Rhr]"] = "OK371-gbb; G85R(cr/hr)"
temp$Genotype[temp$Genotype == "OK371>gbb [G85cr/G85Rhr]"] = "OK371-gbb; G85R(cr/hr)"

temp$Genotype[temp$Genotype == "OK371 > saxA [WT]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371 > saxA [G85R]"] = "OK371-saxA; G85R(cr/hr)"
temp$Genotype[temp$Genotype == "OK371> saxA [WT-cr/LoxP]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>saxA G85R"] = "OK371-saxA; G85R(cr/hr)"
temp$Genotype[temp$Genotype == "OK371>saxA [WTcr/LoxP]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>saxA [WT-cr/LoxP]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>saxA [G85Rcr/G85Rhr]"] = "OK371-saxA; G85R(cr/hr)"


temp$Genotype = factor(temp$Genotype, levels = c("WT(cr/LoxP)",
                                                 "OK371-gbb; WT(cr/LoxP)",
                                                 "OK371-saxA; WT(cr/LoxP)",
                                                 "G85R(cr/hr)",
                                                 "OK371-gbb; G85R(cr/hr)",
                                                 "OK371-saxA; G85R(cr/hr)"))

VNC = temp[temp$Data == "VNC",]
VNC$CTCF = VNC$CTCF/1000
geno = unique(VNC$Genotype)

VNC$Genotype = factor(VNC$Genotype, levels = c("WT(cr/LoxP)",
                                                 "OK371-gbb; WT(cr/LoxP)",
                                                 "OK371-saxA; WT(cr/LoxP)",
                                                 "G85R(cr/hr)",
                                                 "OK371-gbb; G85R(cr/hr)",
                                                 "OK371-saxA; G85R(cr/hr)"))


par(mar=c(5.1, 4.1, 4.1, 6.3))

boxplot(VNC$CTCF~VNC$Genotype,
        at = c(1,2,3, 5,6,7),
        xaxt = 'none',
        las = 2,
        xlab = "",
        col = hex(c("grey", "steelblue", "firebrick"), .25),
        ylab = "RFU",
        main = "Larval A3 MN nuclei pMad fluorescence",
        ylim = c(0,1800))


beeswarm(VNC$CTCF~VNC$Genotype,
         at = c(1,2,3, 5,6,7),
       pch = 21,
       cex = 1.0,
       spacing = 1.0,
       method = "swarm",
       bg = c("grey", "steelblue", "firebrick"),
       add = T)


axis(side = 1,
     at = c(2, 6),
     labels = c("WT(cr/hr)", "G85R(cr/hr)"))

legend('topright',
       inset=c(-0.225,0),
       legend = c('Control', 'gbb', "SaxA"), 
       fill = c("grey", "steelblue", "firebrick"),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

lines(x=c(1,5),
      y=c(1500,1500))

lines(x=c(5,7),
      y=c(1600,1600))

text(x = 3,
     y = 1550,
     labels = "*")

text(x = 6,
     y = 1650,
     labels = "*")


abline(v = 4, lty = 2)

```

**Figure 3D** Total pMad fluorescence was modeled as a function of genotype, rescue, and their interaction. The model did not have a significant main effect alone but the interaction (p = 0.0445) and rescue (p = 0.02737) effects were significant. Post hoc pairwise analysis at each genotype*rescue was done by comparing the estimated marginal means (emmeans) and found that there was a significant difference in total pMad fluorescence between the comparisons WT(cr/hr) - G85R(cr/hr) (p = 0.0211) and OK371>SaxA, G85R - G85R(cr/hr) (p = 0.0177). <a href="https://docs.google.com/spreadsheets/d/18RMFSDqxhagL9BhQNR42RdEYw_tSqaWRmshrWkc8Tig/edit?usp=sharing" target="_blank">Fig3D Data</a><br><br>

```{r echo =F, warning=F, include = F}

VNC$Rescue = "Control"
VNC$Rescue[grep(pattern = "gbb", VNC$Genotype)] = "gbb"
VNC$Rescue[grep(pattern = "saxA", VNC$Genotype)] = "saxA"
VNC$Model = "WT"
VNC$Model[grep(pattern = "G85R", VNC$Genotype)] = "G85R"

model <- lm(CTCF ~ Rescue * Model + Experiment, data=VNC)
anova(model)

emmeans(model, pairwise ~ Model | Rescue)

emmeans(model, pairwise ~ Rescue | Model)

print("Pairwise comparisons were performed using estimated marginal means derived from a linear model, with Tukey-adjusted p-values.")

```

<a id="fig4a"></a>

## Figure 4A: A4V Female ChAT>gbb Climbing

```{r include = F, echo = F}

fig4ab = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 4AB.csv", header = T)


fig4ab$sex = "F"
fig4ab$sex[grep("_m_", fig4ab$vial_ID)] = "M"
fig4ab$sample = paste0(fig4ab$geno, fig4ab$sex, fig4ab$day)
#fig4ab = fig4ab[!(fig4ab$sex == "F" & fig4ab$day == 2),]

fig4ab.mean = unique(fig4ab[,c("geno", "sex", "day", "sample")])
fig4ab.mean$mean = 0
i=1
while(i<=nrow(fig4ab.mean)){
  temp = na.omit(subset(fig4ab, sample == fig4ab.mean$sample[i]))
  fig4ab.mean$mean[i] = mean(temp$corrected.values)
  i=i+1
}


fig4cd = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 4CD.csv", header = T)

fig4cd = fig4cd[fig4cd$day != 9,]

fig4cd$sex = "F"
fig4cd$sex[grep("_m_", fig4cd$vial_ID)] = "M"
fig4cd$sample = paste0(fig4cd$geno, fig4cd$sex, fig4cd$day)

fig4cd.mean = unique(fig4cd[,c("geno", "sex", "day", "sample")])
fig4cd.mean$mean = 0
i=1
while(i<=nrow(fig4cd.mean)){
  temp = na.omit(subset(fig4cd, sample == fig4cd.mean$sample[i]))
  fig4cd.mean$mean[i] = mean(temp$corrected.values)
  i=i+1
}

fgbb = subset(fig4ab, sex == "F")
fchagbb = fgbb[grep("ok", fgbb$geno, invert = T),]
fokgbb = fgbb[grep("cha", fgbb$geno, invert = T),]

mgbb = subset(fig4ab, sex == "M")
mchagbb = mgbb[grep("ok", mgbb$geno, invert = T),]
mokgbb = mgbb[grep("cha", mgbb$geno, invert = T),]

fsax = subset(fig4cd, sex == "F")
fchasax = fsax[grep("ok", fsax$geno, invert = T),]
foksax = fsax[grep("cha", fsax$geno, invert = T),]

msax = subset(fig4ab, sex == "M")
mchasax = msax[grep("ok", msax$geno, invert = T),]
moksax = msax[grep("cha", msax$geno, invert = T),]

pcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")


```

```{r echo = F, fig.width=10, fig.height=6, fig4a}

f.mean = subset(fig4ab.mean, sex == "F")
f.mean = f.mean[order(f.mean$day),]

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 3.75),
     xlim = c(1,27),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Females",
     frame.plot = F)

axis(1, at = 1:27, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:13), tick = F, line = -0.75)

axis(2, at = 0:3, label = F, lwd = 3, line = -0.05)
axis(2, at = 0:3, tick = F, las = 1, line = -0.05)

lines(x=c(0,0),
      y=c(0,3.5),
      lwd = 3,
      xpd = T)

lines(x=c(0,27),
      y=c(0,0),
      lwd = 3,
      xpd = T)


lines(x = f.mean[f.mean$geno == 'silent', 'day'],
      y = f.mean[f.mean$geno == 'silent', 'mean'],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = f.mean[f.mean$geno == 'A4V', 'day'],
      y = f.mean[f.mean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = f.mean[f.mean$geno == 'silentcha', 'day'],
      y = f.mean[f.mean$geno == 'silentcha', 'mean'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = f.mean[f.mean$geno == 'A4Vcha', 'day'],
      y = f.mean[f.mean$geno == 'A4Vcha', 'mean'],
      type = 'l',
      col = pcol[4],
      lwd = 2)



points(x = f.mean[f.mean$geno == 'silent', 'day'],
       y = f.mean[f.mean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4V', 'day'],
       y = f.mean[f.mean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)
points(x = f.mean[f.mean$geno == 'silentcha', 'day'],
       y = f.mean[f.mean$geno == 'silentcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4Vcha', 'day'],
       y = f.mean[f.mean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)




points(x = f.mean[f.mean$geno == 'silent', 'day'],
       y = f.mean[f.mean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex = 1)

points(x = f.mean[f.mean$geno == 'A4V', 'day'],
       y = f.mean[f.mean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex = 1)
points(x = f.mean[f.mean$geno == 'silentcha', 'day'],
       y = f.mean[f.mean$geno == 'silentcha', 'mean'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4Vcha', 'day'],
       y = f.mean[f.mean$geno == 'A4Vcha', 'mean'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex = 1)

legend("topright",
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V'),
       pch = c(23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5),
                hex(pcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```

**Figure 4A** Climbing plots show the mean at each time point. Climbing rate was fit to a linear mixed-effects model to account for genotype, age, and repeated measurements from the same vial. Model term significance was then analyzed using an ANOVA. There was a significant genotype (p = 1.882e-11), day (p = < 2.2e-16) and the interaction effect (p = 1.954e-05) on climbing rate. Post hoc pairwise analysis of the significant interaction was done using estimated marginal means (emmeans) and comparison of the genotype specific slopes (emtrends). Comparing the genotype specific slopes found that ChAT>gbb, A4V had a significantly faster decrease in activity over time than A4V (p = <.0001), Silent (p = 0.0010), and ChAT>gbb, Silent (p = 0.0469). The estimated marginal means were found to be significantly lower for A4V than ChAT>gbb, A4V (p = <.0001), Silent (p = <.0001), and ChAT>gbb, Silent (<.0001). <a href="https://docs.google.com/spreadsheets/d/1zjq2tJQWOyP37nqseViUoOh9V_PmVfp-HCHkWna558E/edit?usp=sharing" target="_blank">Fig4ABCD Data</a><br><br>

```{r echo = F, include = F}

##lmer = linear mixed-effect model
##interaction effect (geno*day) for each genotype, generate a model that estimates velocity at a given time. Treats geno as a factor and time as a number due to the type of data it is (factor and numeric). Change day to factor if you don't want to assume a steady decline or compare rate of decline. That would see if any days are different.

## random effect (1|vial_ID) treats vial_ID as a source of variation that we are not concerned with testing but is potentially adding noise to the model


##the (1 | vial_ID) argument allows the model to treat each vial independently for comparison. For example, if a vial is inherently faster, it doesn't throw off the model for the whole genotype.

##Since day is numeric, the model is testing aging rate instead of determining if genotypes are different at specific days.


fchagbb$geno = factor(fchagbb$geno, levels = c("silent", "silentcha", "A4V", "A4Vcha"))

fchagbb$vial_ID = paste0(fchagbb$geno, substr(fchagbb$vial_ID, nchar(fchagbb$vial_ID), nchar(fchagbb$vial_ID)))

fit <- lmer(
  corrected.values ~ geno * day +
    (1 | vial_ID),
  data = fchagbb
)

anova(fit)
print("anova results")
print("geno-There is a significant difference between genotypes")
print("day-climbing changes over time")
print("geno:day-There is a significant difference in how genotypes change over time")


#summary(fit)

#emmeans(fit, ~ geno)                    # overall levels
#emtrends(fit, ~ geno, var="day")        # aging rates
#emmeans(fit, ~ geno | day,
#        at=list(day=unique(fchagbb$day)))  # specific timepoints


emm <- emmeans(fit, ~ geno)
pairs(emm, adjust = "holm")
print("calculates model estimated means of each genotype then compares them with a t-test on model adjusted means with holm multiple testing correction")
##estimates the difference in level between 2 conditions and tests if it is significant
print("A4V has significantly lower velocity than each other condition")


trends <- emtrends(fit, ~ geno, var = "day")
pairs(trends, adjust = "holm")
print("calculates model estimated slope for each genotype over time then compares them with a t-test using model adjusted means and estimated standard error with a holm multiple testing correction")
##estimates the difference in age associated change in rate
print("A4V ChAT>gbb has a significantly higher rate of decreasing activity")


##Overall stat summary: I fit a linear mixed-effects model to account for genotype, age, and repeated measurements from the same vial. I then used ANOVA F-tests to assess the significance of the model terms, followed by post hoc pairwise t-tests on model-estimated marginal means (or slopes) with multiple-testing correction.
```

<a id="fig4b"></a>

## Figure 4B: A4V Female OK371>gbb Climbing

```{r echo = F, fig.width=10, fig.height=6, fig4b}

f.mean = subset(fig4ab.mean, sex == "F")
f.mean = f.mean[order(f.mean$day),]

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 3.75),
     xlim = c(1,27),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Females",
     frame.plot = F)

axis(1, at = 1:27, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:13), tick = F, line = -0.75)

axis(2, at = 0:3, label = F, lwd = 3, line = -0.05)
axis(2, at = 0:3, tick = F, las = 1, line = -0.05)

lines(x=c(0,0),
      y=c(0,3.5),
      lwd = 3,
      xpd = T)

lines(x=c(0,27),
      y=c(0,0),
      lwd = 3,
      xpd = T)


lines(x = f.mean[f.mean$geno == 'silent', 'day'],
      y = f.mean[f.mean$geno == 'silent', 'mean'],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = f.mean[f.mean$geno == 'A4V', 'day'],
      y = f.mean[f.mean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = f.mean[f.mean$geno == 'silentok', 'day'],
      y = f.mean[f.mean$geno == 'silentok', 'mean'],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = f.mean[f.mean$geno == 'A4Vok', 'day'],
      y = f.mean[f.mean$geno == 'A4Vok', 'mean'],
      type = 'l',
      col = pcol[6],
      lwd = 2)



points(x = f.mean[f.mean$geno == 'silent', 'day'],
       y = f.mean[f.mean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4V', 'day'],
       y = f.mean[f.mean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)
points(x = f.mean[f.mean$geno == 'silentok', 'day'],
       y = f.mean[f.mean$geno == 'silentok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4Vok', 'day'],
       y = f.mean[f.mean$geno == 'A4Vok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)




points(x = f.mean[f.mean$geno == 'silent', 'day'],
       y = f.mean[f.mean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex = 1)

points(x = f.mean[f.mean$geno == 'A4V', 'day'],
       y = f.mean[f.mean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex = 1)
points(x = f.mean[f.mean$geno == 'silentok', 'day'],
       y = f.mean[f.mean$geno == 'silentok', 'mean'],
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4Vok', 'day'],
       y = f.mean[f.mean$geno == 'A4Vok', 'mean'],
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex = 1)


legend("topright",
       legend = c('Silent', 'A4V',
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'),
       pch = c(23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[5], .5),
                hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, include=F}

fokgbb$geno = factor(fokgbb$geno, levels = c("silent", "silentok", "A4V", "A4Vok"))

fokgbb$vial_ID = paste0(fokgbb$geno, substr(fokgbb$vial_ID, nchar(fokgbb$vial_ID), nchar(fokgbb$vial_ID)))

fit <- lmer(
  corrected.values ~ geno * day +
    (1 | vial_ID),
  data = fokgbb
)

anova(fit)
print("anova results")
print("geno-There is a significant difference between genotypes")
print("day-climbing changes over time")
print("geno:day-There is NOT a significant difference in how genotypes change over time")

emm <- emmeans(fit, ~ geno)
pairs(emm, adjust = "holm")
print("calculates model estimated means of each genotype then compares them with a t-test on model adjusted means with holm multiple testing correction")
print("Almost all comparisons were significantly different. Specifically, A4V has significantly lower velocity than each other condition and A4V- OK371>gbb is significantly lower than Silent-OK371>gbb. ")


trends <- emtrends(fit, ~ geno, var = "day")
pairs(trends, adjust = "holm")
print("calculates model estimated slope for each genotype over time then compares them with a t-test using model adjusted means and estimated standard error with a holm multiple testing correction")
##estimates the difference in age associated change in rate
print("None of the genotypes have a significantly different rate of decreasing activity")

```

**Figure 4B** Climbing plots show the mean at each time point. Climbing rate was fit to a linear mixed-effects model to account for genotype, age, and repeated measurements from the same vial. Model term significance was then analyzed using an ANOVA. There was a significant genotype (p = 5.907e-14) and day (p = < 2.2e-16) effect on climbing rate, but the interaction effect was not significant. Post hoc pairwise analysis of the significant interaction was done using estimated marginal means (emmeans) and comparison of the genotype specific slopes (emtrends). Comparing the genotype specific slopes found that none of the rates of decline were significantly different for the genotypes. The estimated marginal means were found to be significantly lower for A4V than ChAT>gbb, A4V (p = <.0001), Silent (p = <.0001), and ChAT>gbb, Silent (<.0001).<br><br>

<a id="fig4c"></a>

## Figure 4C: A4V Male ChAT>gbb Climbing

```{r echo = F, fig.width=10, fig.height=6, fig4c}

m.mean = subset(fig4ab.mean, sex == "M")
m.mean = m.mean[order(m.mean$day),]

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 3.75),
     xlim = c(1,27),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Males",
     frame.plot = F)

axis(1, at = 1:27, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:13), tick = F, line = -0.75)

axis(2, at = 0:3, label = F, lwd = 3, line = -0.05)
axis(2, at = 0:3, tick = F, las = 1, line = -0.05)

lines(x=c(0,0),
      y=c(0,3.5),
      lwd = 3,
      xpd = T)

lines(x=c(0,27),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = m.mean[m.mean$geno == 'silent', 'day'],
      y = m.mean[m.mean$geno == 'silent', 'mean'],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'A4V', 'day'],
      y = m.mean[m.mean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'silentcha', 'day'],
      y = m.mean[m.mean$geno == 'silentcha', 'mean'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'A4Vcha', 'day'],
      y = m.mean[m.mean$geno == 'A4Vcha', 'mean'],
      type = 'l',
      col = pcol[4],
      lwd = 2)

points(x = m.mean[m.mean$geno == 'silent', 'day'],
       y = m.mean[m.mean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4V', 'day'],
       y = m.mean[m.mean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)
points(x = m.mean[m.mean$geno == 'silentcha', 'day'],
       y = m.mean[m.mean$geno == 'silentcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4Vcha', 'day'],
       y = m.mean[m.mean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)




points(x = m.mean[m.mean$geno == 'silent', 'day'],
       y = m.mean[m.mean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex = 1)

points(x = m.mean[m.mean$geno == 'A4V', 'day'],
       y = m.mean[m.mean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex = 1)
points(x = m.mean[m.mean$geno == 'silentcha', 'day'],
       y = m.mean[m.mean$geno == 'silentcha', 'mean'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4Vcha', 'day'],
       y = m.mean[m.mean$geno == 'A4Vcha', 'mean'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex = 1)

legend("topright",
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5), hex(pcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, include = F}

mchagbb$geno = factor(mchagbb$geno, levels = c("silent", "silentcha", "A4V", "A4Vcha"))

mchagbb$vial_ID = paste0(mchagbb$geno, substr(mchagbb$vial_ID, nchar(mchagbb$vial_ID), nchar(mchagbb$vial_ID)))

fit <- lmer(
  corrected.values ~ geno * day +
    (1 | vial_ID),
  data = mchagbb
)

anova(fit)
print("anova results")
print("geno-There is a significant difference between genotypes")
print("day-climbing changes over time")
print("geno:day-There is  a significant difference in how genotypes change over time")

emm <- emmeans(fit, ~ geno)
pairs(emm, adjust = "holm")
print("calculates model estimated means of each genotype then compares them with a t-test on model adjusted means with holm multiple testing correction")
print("A4V has significantly lower velocity than each other condition. ")


trends <- emtrends(fit, ~ geno, var = "day")
pairs(trends, adjust = "holm")
print("calculates model estimated slope for each genotype over time then compares them with a t-test using model adjusted means and estimated standard error with a holm multiple testing correction")
##estimates the difference in age associated change in rate
print("A4V activity decreases at a significantly faster rate than all the other genotypes")

```

**Figure 4C** Climbing plots show the mean at each time point. Climbing rate was fit to a linear mixed-effects model to account for genotype, age, and repeated measurements from the same vial. Model term significance was then analyzed using an ANOVA. There was a significant genotype (p = 0.0001079), day (p = < 2.2e-16) and the interaction effect (p = 0.0074109) on climbing rate. Post hoc pairwise analysis of the significant interaction was done using estimated marginal means (emmeans) and comparison of the genotype specific slopes (emtrends). Comparing the genotype specific slopes found that A4V had a significantly faster decrease in activity over time than ChAT>gbb, A4V (p = 0.0189), Silent (p = 0.0069), and ChAT>gbb, Silent (0.0082). The estimated marginal means were found to be significantly lower for A4V than ChAT>gbb, A4V (p = <.0001), Silent (p = <.0001), and ChAT>gbb, Silent (<.0001).<br><br>

<a id="fig4d"></a>

## Figure 4D: A4V Male OK371>gbb Climbing

```{r echo = F, fig.width=10, fig.height=6, fig4d}

m.mean = subset(fig4ab.mean, sex == "M")
m.mean = m.mean[order(m.mean$day),]

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 3.75),
     xlim = c(1,27),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Males",
     frame.plot = F)

axis(1, at = 1:27, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:13), tick = F, line = -0.75)

axis(2, at = 0:3, label = F, lwd = 3, line = -0.05)
axis(2, at = 0:3, tick = F, las = 1, line = -0.05)

lines(x=c(0,0),
      y=c(0,3.5),
      lwd = 3,
      xpd = T)

lines(x=c(0,27),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = m.mean[m.mean$geno == 'silent', 'day'],
      y = m.mean[m.mean$geno == 'silent', 'mean'],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'A4V', 'day'],
      y = m.mean[m.mean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'silentok', 'day'],
      y = m.mean[m.mean$geno == 'silentok', 'mean'],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'A4Vok', 'day'],
      y = m.mean[m.mean$geno == 'A4Vok', 'mean'],
      type = 'l',
      col = pcol[6],
      lwd = 2)

points(x = m.mean[m.mean$geno == 'silent', 'day'],
       y = m.mean[m.mean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4V', 'day'],
       y = m.mean[m.mean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)
points(x = m.mean[m.mean$geno == 'silentok', 'day'],
       y = m.mean[m.mean$geno == 'silentok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4Vcha', 'day'],
       y = m.mean[m.mean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)


points(x = m.mean[m.mean$geno == 'silent', 'day'],
       y = m.mean[m.mean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex = 1)

points(x = m.mean[m.mean$geno == 'A4V', 'day'],
       y = m.mean[m.mean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex = 1)
points(x = m.mean[m.mean$geno == 'silentok', 'day'],
       y = m.mean[m.mean$geno == 'silentok', 'mean'],
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4Vok', 'day'],
       y = m.mean[m.mean$geno == 'A4Vok', 'mean'],
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex = 1)

legend("topright",
       legend = c('Silent', 'A4V',
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[5], .5), hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, include = F}

mokgbb$geno = factor(mokgbb$geno, levels = c("silent", "silentok", "A4V", "A4Vok"))

mokgbb$vial_ID = paste0(mokgbb$geno, substr(mokgbb$vial_ID, nchar(mokgbb$vial_ID), nchar(mokgbb$vial_ID)))

fit <- lmer(
  corrected.values ~ geno * day +
    (1 | vial_ID),
  data = mokgbb
)

anova(fit)
print("anova results")
print("geno-There is a significant difference between genotypes")
print("day-climbing changes over time")
print("geno:day-There is  a significant difference in how genotypes change over time")

emm <- emmeans(fit, ~ geno)
pairs(emm, adjust = "holm")
print("calculates model estimated means of each genotype then compares them with a t-test on model adjusted means with holm multiple testing correction")
print("A4V has significantly lower velocity than each other condition. ")


trends <- emtrends(fit, ~ geno, var = "day")
pairs(trends, adjust = "holm")
print("calculates model estimated slope for each genotype over time then compares them with a t-test using model adjusted means and estimated standard error with a holm multiple testing correction")
##estimates the difference in age associated change in rate
print("All genotype comparisons are significant, but the A4V-OK371>gbb compared to Silent-OK371>gbb and to A4V were the most significant. Important note, since we are comparing slopes and taking the vial in to account as a factor, ")

emtrends(fit, ~ geno, var = "day")
print("The estimated rate of declines (day.trend). The slowest rate of decline is A4Vok, followed by Silent, then Silentok with A4V having the fastest estimated rate of decline. Notably, the SE was distinctly higher in A4V explaining the observed p.value.")
```

**Figure 4D** Climbing plots show the mean at each time point. Climbing rate was fit to a linear mixed-effects model to account for genotype, age, and repeated measurements from the same vial. Model term significance was then analyzed using an ANOVA. There was a significant genotype (p = 1.289e-07), day (p = < 2.2e-16) and the interaction effect (p = 2.112e-07) on climbing rate. Post hoc pairwise analysis of the significant interaction was done using estimated marginal means (emmeans) and comparison of the genotype specific slopes (emtrends). Comparing the genotype specific slopes found that A4V had a significantly faster decrease in activity over time than OK371>gbb, A4V (p = 0.0014), Silent (p = 0.0117), and OK371>gbb, Silent (p = 0.0428). The estimated marginal means were found to be significantly lower for A4V than OK371>gbb, A4V (p = <.0001), Silent (p = <.0001), and OK371>gbb, Silent (<.0001).<br><br>

<a id="fig4e"></a>

## Figure 4E: A4V Female ChAT>SaxA Climbing

```{r echo = F, fig.width=10, fig.height=6, fig4e}

f.mean = subset(fig4cd.mean, sex == "F")
f.mean = f.mean[order(f.mean$day),]

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 2.25),
     xlim = c(1,32),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Females",
     frame.plot = F)

axis(1, at = 1:32, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:16), tick = F, line = -0.75)

axis(2, at = 0:2, label = F, lwd = 3, line = -0.2)
axis(2, at = 0:2, tick = F, las = 1, line = -0.2)

lines(x=c(0,0),
      y=c(0,2),
      lwd = 3,
      xpd = T)

lines(x=c(0,32),
      y=c(0,0),
      lwd = 3,
      xpd = T)


lines(x = f.mean[f.mean$geno == 'silent', 'day'],
      y = f.mean[f.mean$geno == 'silent', 'mean'],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = f.mean[f.mean$geno == 'A4V', 'day'],
      y = f.mean[f.mean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = f.mean[f.mean$geno == 'silentcha', 'day'],
      y = f.mean[f.mean$geno == 'silentcha', 'mean'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = f.mean[f.mean$geno == 'A4Vcha', 'day'],
      y = f.mean[f.mean$geno == 'A4Vcha', 'mean'],
      type = 'l',
      col = pcol[4],
      lwd = 2)

points(x = f.mean[f.mean$geno == 'silent', 'day'],
       y = f.mean[f.mean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4V', 'day'],
       y = f.mean[f.mean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)
points(x = f.mean[f.mean$geno == 'silentcha', 'day'],
       y = f.mean[f.mean$geno == 'silentcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4Vcha', 'day'],
       y = f.mean[f.mean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)

points(x = f.mean[f.mean$geno == 'silent', 'day'],
       y = f.mean[f.mean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex = 1)

points(x = f.mean[f.mean$geno == 'A4V', 'day'],
       y = f.mean[f.mean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex = 1)
points(x = f.mean[f.mean$geno == 'silentcha', 'day'],
       y = f.mean[f.mean$geno == 'silentcha', 'mean'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4Vcha', 'day'],
       y = f.mean[f.mean$geno == 'A4Vcha', 'mean'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex = 1)

legend("topright",
       legend = c('Silent', 'A4V',
                  'ChAT>saxA;Silent', 'ChAT>saxA;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5), hex(pcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```

**Figure 4E** Climbing plots show the mean at each time point. Climbing rate was fit to a linear mixed-effects model to account for genotype, age, and repeated measurements from the same vial. Model term significance was then analyzed using an ANOVA. There was a significant genotype (p = < 2.2e-16), day (p = < 2.2e-16) and the interaction effect (p = 2.112e-07) on climbing rate. Post hoc pairwise analysis of the significant interaction was done using estimated marginal means (emmeans) and comparison of the genotype specific slopes (emtrends). Comparing the genotype specific slopes found that A4V had a significantly slower decrease in activity over time than ChAT>SaxA, A4V (p = <.0001), Silent (p = <.0001), and ChAT>SaxA, Silent (<.0001). The estimated marginal means were found to be significantly lower for A4V than ChAT>SaxA, A4V (p = <.0001), Silent (p = <.0001), and ChAT>SaxA, Silent (<.0001). <a href="https://docs.google.com/spreadsheets/d/1DFGjCQngjJ6EP_f6i5D3gnMSfJ2tiLoK_QyIkrhGc2s/edit?usp=sharing" target="_blank">Fig4EFGH Data</a><br><br>

```{r echo = F, include = F}

fchasax$geno = factor(fchasax$geno, levels = c("silent", "silentcha", "A4V", "A4Vcha"))

fchasax$vial_ID = paste0(fchasax$geno, substr(fchasax$vial_ID, nchar(fchasax$vial_ID), nchar(fchasax$vial_ID)))

fit <- lmer(
  corrected.values ~ geno * day +
    (1 | vial_ID),
  data = fchasax
)

anova(fit)
print("anova results")
print("geno-There is a significant difference between genotypes")
print("day-climbing changes over time")
print("geno:day-There is  a significant difference in how genotypes change over time")

emm <- emmeans(fit, ~ geno)
pairs(emm, adjust = "holm")
print("calculates model estimated means of each genotype then compares them with a t-test on model adjusted means with holm multiple testing correction")
print("A4V has significantly lower velocity than each other condition. ")


trends <- emtrends(fit, ~ geno, var = "day")
pairs(trends, adjust = "holm")
print("calculates model estimated slope for each genotype over time then compares them with a t-test using model adjusted means and estimated standard error with a holm multiple testing correction")
##estimates the difference in age associated change in rate
print("A4V has a significantly lower rate of decline than all other genotypes.")

emtrends(fit, ~ geno, var = "day")
print("The estimated rate of declines (day.trend). The slowest rate of decline by far is is A4V.")
```

<a id="fig4f"></a>

## Figure 4F: A4V Female OK371>SaxA Climbing

```{r echo = F, fig.width=10, fig.height=6, fig4f}

f.mean = subset(fig4cd.mean, sex == "F")
f.mean = f.mean[order(f.mean$day),]

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 2.25),
     xlim = c(1,32),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Females",
     frame.plot = F)

axis(1, at = 1:32, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:16), tick = F, line = -0.75)

axis(2, at = 0:2, label = F, lwd = 3, line = -0.2)
axis(2, at = 0:2, tick = F, las = 1, line = -0.2)

lines(x=c(0,0),
      y=c(0,2),
      lwd = 3,
      xpd = T)

lines(x=c(0,32),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = f.mean[f.mean$geno == 'silent', 'day'],
      y = f.mean[f.mean$geno == 'silent', 'mean'],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = f.mean[f.mean$geno == 'A4V', 'day'],
      y = f.mean[f.mean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)



lines(x = f.mean[f.mean$geno == 'silentok', 'day'],
      y = f.mean[f.mean$geno == 'silentok', 'mean'],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = f.mean[f.mean$geno == 'A4Vok', 'day'],
      y = f.mean[f.mean$geno == 'A4Vok', 'mean'],
      type = 'l',
      col = pcol[6],
      lwd = 2)

points(x = f.mean[f.mean$geno == 'silent', 'day'],
       y = f.mean[f.mean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4V', 'day'],
       y = f.mean[f.mean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)

points(x = f.mean[f.mean$geno == 'silentok', 'day'],
       y = f.mean[f.mean$geno == 'silentok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4Vok', 'day'],
       y = f.mean[f.mean$geno == 'A4Vok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)

points(x = f.mean[f.mean$geno == 'silent', 'day'],
       y = f.mean[f.mean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex = 1)

points(x = f.mean[f.mean$geno == 'A4V', 'day'],
       y = f.mean[f.mean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex = 1)

points(x = f.mean[f.mean$geno == 'silentok', 'day'],
       y = f.mean[f.mean$geno == 'silentok', 'mean'],
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex = 1)
points(x = f.mean[f.mean$geno == 'A4Vok', 'day'],
       y = f.mean[f.mean$geno == 'A4Vok', 'mean'],
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex = 1)

legend("topright",
       legend = c('Silent', 'A4V',
                  'OK371>saxA;Silent', 'OK371>saxA;A4V'),
       pch = c(23,21,23,21,23,21, 26, 32),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[5], .5), hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```

**Figure 4F** Climbing plots show the mean at each time point. Climbing rate was fit to a linear mixed-effects model to account for genotype, age, and repeated measurements from the same vial. Model term significance was then analyzed using an ANOVA. There was a significant genotype (p = < 2.2e-16), day (p = < 2.2e-16) and the interaction effect (p = 4.809e-11) on climbing rate. Post hoc pairwise analysis of the significant interaction was done using estimated marginal means (emmeans) and comparison of the genotype specific slopes (emtrends). Comparing the genotype specific slopes found that A4V had a significantly slower decrease in activity over time than OK371>SaxA, A4V (p = <.0001), Silent (p = <.0001), and OK371>SaxA, Silent (p <.0001). The estimated marginal means were found to be significantly lower for A4V than OK371>SaxA, A4V (p = <.0001), Silent (p = <.0001), and OK371>SaxA, Silent (<.0001).<br><br>

```{r echo = F, include = F}

foksax$geno = factor(foksax$geno, levels = c("silent", "silentok", "A4V", "A4Vok"))

foksax$vial_ID = paste0(foksax$geno, substr(foksax$vial_ID, nchar(foksax$vial_ID), nchar(foksax$vial_ID)))

fit <- lmer(
  corrected.values ~ geno * day +
    (1 | vial_ID),
  data = foksax
)

anova(fit)
print("anova results")
print("geno-There is a significant difference between genotypes")
print("day-climbing changes over time")
print("geno:day-There is  a significant difference in how genotypes change over time")

emm <- emmeans(fit, ~ geno)
pairs(emm, adjust = "holm")
print("calculates model estimated means of each genotype then compares them with a t-test on model adjusted means with holm multiple testing correction")
print("A4V has significantly lower velocity than each other condition. ")


trends <- emtrends(fit, ~ geno, var = "day")
pairs(trends, adjust = "holm")
print("calculates model estimated slope for each genotype over time then compares them with a t-test using model adjusted means and estimated standard error with a holm multiple testing correction")
##estimates the difference in age associated change in rate
print("A4V has a significantly lower rate of decline than all other genotypes.")

emtrends(fit, ~ geno, var = "day")
print("The estimated rate of declines (day.trend). The slowest rate of decline by far is is A4V.")
```

<a id="fig4g"></a>

## Figure 4G: A4V Male ChAT>SaxA Climbing

```{r echo = F, fig.width=10, fig.height=6, fig4g}

m.mean = subset(fig4cd.mean, sex == "M")
m.mean = m.mean[order(m.mean$day),]

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 2.25),
     xlim = c(1,32),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Males",
     frame.plot = F)

axis(1, at = 1:32, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:16), tick = F, line = -0.75)

axis(2, at = 0:2, label = F, lwd = 3, line = -0.2)
axis(2, at = 0:2, tick = F, las = 1, line = -0.2)

lines(x=c(0,0),
      y=c(0,2),
      lwd = 3,
      xpd = T)

lines(x=c(0,32),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = m.mean[m.mean$geno == 'silent', 'day'],
      y = m.mean[m.mean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'A4V', 'day'],
      y = m.mean[m.mean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'silentcha', 'day'],
      y = m.mean[m.mean$geno == 'silentcha', 'mean'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'A4Vcha', 'day'],
      y = m.mean[m.mean$geno == 'A4Vcha', 'mean'],
      type = 'l',
      col = pcol[4],
      lwd = 2)

points(x = m.mean[m.mean$geno == 'silent', 'day'],
       y = m.mean[m.mean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4V', 'day'],
       y = m.mean[m.mean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)
points(x = m.mean[m.mean$geno == 'silentcha', 'day'],
       y = m.mean[m.mean$geno == 'silentcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4Vcha', 'day'],
       y = m.mean[m.mean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)

points(x = m.mean[m.mean$geno == 'silent', 'day'],
       y = m.mean[m.mean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex = 1)

points(x = m.mean[m.mean$geno == 'A4V', 'day'],
       y = m.mean[m.mean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex = 1)
points(x = m.mean[m.mean$geno == 'silentcha', 'day'],
       y = m.mean[m.mean$geno == 'silentcha', 'mean'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4Vcha', 'day'],
       y = m.mean[m.mean$geno == 'A4Vcha', 'mean'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex = 1)

legend("topright",
       legend = c('Silent', 'A4V',
                  'ChAT>saxA;Silent', 'ChAT>saxA;A4V'),
       pch = c(23,21,23,21,23,21, 26, 32),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5),
                hex(pcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, include = F}

mchasax$geno = factor(mchasax$geno, levels = c("silent", "silentcha", "A4V", "A4Vcha"))

mchasax$vial_ID = paste0(mchasax$geno, substr(mchasax$vial_ID, nchar(mchasax$vial_ID), nchar(mchasax$vial_ID)))

fit <- lmer(
  corrected.values ~ geno * day +
    (1 | vial_ID),
  data = mchasax
)

anova(fit)
print("anova results")
print("geno-There is a significant difference between genotypes")
print("day-climbing changes over time")
print("geno:day-There is  a significant difference in how genotypes change over time")

emm <- emmeans(fit, ~ geno)
pairs(emm, adjust = "holm")
print("calculates model estimated means of each genotype then compares them with a t-test on model adjusted means with holm multiple testing correction")
print("A4V has significantly lower velocity than each other condition. ")


trends <- emtrends(fit, ~ geno, var = "day")
pairs(trends, adjust = "holm")
print("calculates model estimated slope for each genotype over time then compares them with a t-test using model adjusted means and estimated standard error with a holm multiple testing correction")
##estimates the difference in age associated change in rate
print("A4V has a significantly higher rate of decline than all other genotypes.")

emtrends(fit, ~ geno, var = "day")
print("The estimated rate of declines (day.trend). The highest rate of decline by far is A4V which is almost double that of all other genotypes even though it's only 2 days.")
```

**Figure 4G** Climbing plots show the mean at each time point. Climbing rate was fit to a linear mixed-effects model to account for genotype, age, and repeated measurements from the same vial. Model term significance was then analyzed using an ANOVA. There was a significant genotype (p = < 0.0001079), day (p = < 2.2e-16) and the interaction effect (p = 0.0074109) on climbing rate. Post hoc pairwise analysis of the significant interaction was done using estimated marginal means (emmeans) and comparison of the genotype specific slopes (emtrends). Comparing the genotype specific slopes found that A4V had a significantly slower decrease in activity over time than ChAT>SaxA, A4V (p = 0.0189), Silent (p = 0.0082), and ChAT>SaxA, Silent ( p = 0.0069). The estimated marginal means were found to be significantly lower for A4V than ChAT>SaxA, A4V (p = <.0001), Silent (p = <.0001), and ChAT>SaxA, Silent (<.0001).<br><br>

<a id="fig4h"></a>

## Figure 4H: A4V Male OK371>SaxA Climbing

```{r echo = F, fig.width=10, fig.height=6, fig4h}

m.mean = subset(fig4cd.mean, sex == "M")
m.mean = m.mean[order(m.mean$day),]

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 2.25),
     xlim = c(1,32),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Males",
     frame.plot = F)

axis(1, at = 1:32, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:16), tick = F, line = -0.75)

axis(2, at = 0:2, label = F, lwd = 3, line = -0.2)
axis(2, at = 0:2, tick = F, las = 1, line = -0.2)

lines(x=c(0,0),
      y=c(0,2),
      lwd = 3,
      xpd = T)

lines(x=c(0,32),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = m.mean[m.mean$geno == 'silent', 'day'],
      y = m.mean[m.mean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'A4V', 'day'],
      y = m.mean[m.mean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'silentok', 'day'],
      y = m.mean[m.mean$geno == 'silentok', 'mean'],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = m.mean[m.mean$geno == 'A4Vok', 'day'],
      y = m.mean[m.mean$geno == 'A4Vok', 'mean'],
      type = 'l',
      col = pcol[6],
      lwd = 2)

points(x = m.mean[m.mean$geno == 'silent', 'day'],
       y = m.mean[m.mean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4V', 'day'],
       y = m.mean[m.mean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)
points(x = m.mean[m.mean$geno == 'silentok', 'day'],
       y = m.mean[m.mean$geno == 'silentok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4Vcha', 'day'],
       y = m.mean[m.mean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex = 1)

points(x = m.mean[m.mean$geno == 'silent', 'day'],
       y = m.mean[m.mean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex = 1)

points(x = m.mean[m.mean$geno == 'A4V', 'day'],
       y = m.mean[m.mean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex = 1)
points(x = m.mean[m.mean$geno == 'silentok', 'day'],
       y = m.mean[m.mean$geno == 'silentok', 'mean'],
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex = 1)
points(x = m.mean[m.mean$geno == 'A4Vok', 'day'],
       y = m.mean[m.mean$geno == 'A4Vok', 'mean'],
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex = 1)

legend("topright",
       legend = c('Silent', 'A4V',
                  'OK371>saxA;Silent', 'OK371>saxA;A4V'),
       pch = c(23,21,23,21,23,21, 26, 32),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[5], .5),
                hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, include = F}

moksax$geno = factor(moksax$geno, levels = c("silent", "silentok", "A4V", "A4Vok"))

moksax$vial_ID = paste0(moksax$geno, substr(moksax$vial_ID, nchar(moksax$vial_ID), nchar(moksax$vial_ID)))

fit <- lmer(
  corrected.values ~ geno * day +
    (1 | vial_ID),
  data = moksax
)

anova(fit)
print("anova results")
print("geno-There is a significant difference between genotypes")
print("day-climbing changes over time")
print("geno:day-There is  a significant difference in how genotypes change over time")

emm <- emmeans(fit, ~ geno)
pairs(emm, adjust = "holm")
print("calculates model estimated means of each genotype then compares them with a t-test on model adjusted means with holm multiple testing correction")
print("A4V has significantly lower velocity than each other condition. ")


trends <- emtrends(fit, ~ geno, var = "day")
pairs(trends, adjust = "holm")
print("calculates model estimated slope for each genotype over time then compares them with a t-test using model adjusted means and estimated standard error with a holm multiple testing correction")
##estimates the difference in age associated change in rate
print("A4V has a significantly higher rate of decline than all other genotypes. All comparisons were again significant.")

emtrends(fit, ~ geno, var = "day")
print("The estimated rate of declines (day.trend). The fastest rate of decline by far is is A4V, then silentok, then silent and A4Vok has the slowest rate of decline.")
```

**Figure 4H** Climbing plots show the mean at each time point. Climbing rate was fit to a linear mixed-effects model to account for genotype, age, and repeated measurements from the same vial. Model term significance was then analyzed using an ANOVA. There was a significant genotype (p = < 1.289e-07), day (p = < 2.2e-16) and the interaction effect (p = 2.112e-07) on climbing rate. Post hoc pairwise analysis of the significant interaction was done using estimated marginal means (emmeans) and comparison of the genotype specific slopes (emtrends). Comparing the genotype specific slopes found that A4V had a significantly slower decrease in activity over time than OK371>SaxA, A4V (p = 0.0014), Silent (p = 0.0117), and OK371>SaxA, Silent (p = 0.0428). The estimated marginal means were found to be significantly lower for A4V than OK371>SaxA, A4V (p = <.0001), Silent (p = <.0001), and OK371>SaxA, Silent (<.0001).









