---
title: "Russo Fig1 Replots"
author: "John Santiago"
date: "2025-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo = F}

library(ggplot2)
library(beeswarm)
library(kableExtra)
library(emmeans)

std.colors = setNames(c('navy', "gold2", "turquoise4", "royalblue",
                        'orange', "deeppink", "grey", "green4"),
                      c("w1118", "WT(LoxP)", "G85R(hr)", "WT(cr)",
                        "G85R(cr)", "A4V", "A4V/X16", "A4V/G85R(cr)"))

hex = function(color, alpha){
  rgb2hex = color
  i = 1
  while(i <= length(color)){
    rgbvals = col2rgb(color[i])
    rgb2hex[i] = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
    i = i+1
  }
  return(rgb2hex)
}

```


```{r echo = F}

B = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Russo Fig 1 eclosion.csv")
B.totalflies = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Russo Fig 1 eclosion totalflies.csv")
B.labels = c("w1118", "WTloxP", "G85Rhr",
             "WTcr", "G85Rcr",
             "G85R/X16", "WTcr/X16",
             "A4V", "A4V/X16", "A4V/G85Rcr")


colnames(B) = B.labels

mean.B = setNames(rep(0, ncol(B)), colnames(B))
i=1
while(i<=ncol(B)){
  mean.B[i] = mean(na.omit(B[,i]))
  i=i+1
}

##Comparisons for pval 
#WTloxpvG85Rhr
#WTcrvG85RcrvG85R/X16vWT/X16
#A4VvsWTcrvsA4V/X16vs


```

##G85R and A4V models eclosion data

```{r echo = F, fig.width=10, fig.height=7}


fig1b.data = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Russo Fig 1 eclosion full data.csv")

percent.fig1b = setNames(rep(NA_real_,length(unique(fig1b.data$Genotype))),
                         unique(fig1b.data$Genotype))

#adds up all eclosed and not eclosed across replicates before calculating percent eclosed
i=1
while(i<=length(percent.fig1b)){
  tempEclosed = sum(fig1b.data[fig1b.data$Genotype == names(percent.fig1b)[i], "Eclosed"])
  tempNotEclosed = sum(fig1b.data[fig1b.data$Genotype == names(percent.fig1b)[i], "NotEclosed"])
  tempTotal = tempEclosed+tempNotEclosed
  percent.fig1b[i] = signif(100*(tempEclosed/tempTotal),3)
  i=i+1
}

#plots the total percent eclosed
par(mar = c(9,4,4,2))

barplot(height = percent.fig1b,
        ylim = c(0, 120),
        col = "black",
        las = 2,
        ylab = "Eclosion frequency (%)",
        main = "Eclosion from the pupal case",
        xaxt = "n",
        yaxt = "n")
axis(side = 1,
     at = c(.7 + (c(0:9)*1.2)),
     lwd = 3,
     line = 0,
     labels = B.labels,
     xpd = T,
     las = 2)
axis(side = 2,
     at = c(0,20,40,60,80,100),
     lwd = 3,
     line = 0,
     labels = T,
     xpd = T,
     las = 2)

lines(x = c(-0.5,13.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)


lines(x = c(1.9,3.1),
      y = c(105, 105),
      lwd = 3)

lines(x = c(1.9,1.9),
      y = c(105, 102),
      lwd = 3)


text(x = 3.1,
     y = 109,
    labels = "*",
    cex = 2)


lines(x = c(4.3,11.5),
      y = c(105, 105),
      lwd = 3)

lines(x = c(4.3,4.3),
      y = c(105, 102),
      lwd = 3)


text(x = 5.5,
     y = 109,
    labels = "*",
    cex = 2)

text(x = 6.7,
     y = 109,
    labels = "*",
    cex = 2)

text(x = 9.1,
     y = 109,
    labels = "*",
    cex = 2)

text(x = 11.5,
     y = 109,
    labels = "*",
    cex = 2,
    xpd = t)

```

```{r, echo = F}

#Why Fisher’s exact test is the right choice for the situation:

#Binomial outcome (eclosed vs not) 
#Comparing pairs of genotypes (mutant vs matched control) 
#Unequal sample sizes 
#Extreme values (0% / 100%) 
#Moderate number of replicates (4–9 vials) 
#Clear biological grouping (not all comparisons meaningful) 

#This is exactly the kind of scenario Fisher’s test was designed for.

##Another Fishers Test
rawB = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Russo Fig 1 eclosion Raw Data.csv")


rawB <- aggregate(cbind(eclosed, not.eclosed) ~ genotype, data = rawB, sum)

# total and proportion
rawB$total <- rawB$eclosed + rawB$not.eclosed
rawB$prop  <- rawB$eclosed / rawB$total

comparisons <- list(
  c("WTloxP", "G85Rhr"),
  c("WTcr", "G85Rcr"),
  c("WTcr", "G85R/X16"),
  c("WTcr", "WTcr/X16"),
  c("WTcr", "A4V"),
  c("WTcr", "A4V/X16"),
  c("WTcr", "A4V/G85Rcr")
)

results <- data.frame()

for (pair in comparisons) {
  
  g1 <- pair[1]
  g2 <- pair[2]
  
  d1 <- rawB[rawB$genotype == g1, ]
  d2 <- rawB[rawB$genotype == g2, ]
  
  mat <- matrix(c(
    d1$eclosed, d1$not.eclosed,
    d2$eclosed, d2$not.eclosed
  ), nrow = 2, byrow = TRUE)
  
  ft <- fisher.test(mat)
  
  results <- rbind(results, data.frame(
    Genotype1 = g1,
    n1 = d1$total,
    prop1 = d1$prop,
    Genotype2 = g2,
    n2 = d2$total,
    prop2 = d2$prop,
    diff = d1$prop - d2$prop,
    p.value = ft$p.value
  ))
}

results$padj <- p.adjust(results$p.value, method = "BH")

results$prop1 <- round(results$prop1 * 100, 1)
results$prop2 <- round(results$prop2 * 100, 1)
results$diff  <- round(results$diff * 100, 1)
results$p.value <- signif(results$p.value, 3)
results$padj    <- signif(results$padj, 3)

results1b = results

```

```{r echo = F, include = T}

knitr::kable(results1b,
             caption = 'All fishers exact tests for figure 1B eclosion data.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(results1b))[results$padj <= .05], bold = T, color = "black", background = "gold")

```

##Pharate bouton data

```{r echo = F}

C = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1C pharate_bouton_counts.xlsx - counts.csv")

W1118 = data.frame(Genotype = "W1118",
                   Boutons = na.omit(C$W1118),
                   x = 1)
W1118 = W1118[order(W1118$Boutons),]
W1118$x[c(3)] = .85
W1118$x[c(4)] = 1.15


SILENT = data.frame(Genotype = "SILENT",
                   Boutons = na.omit(C$SILENT),
                   x = 2)
SILENT = SILENT[order(SILENT$Boutons),]
SILENT$x[c(3)] = 1.85
SILENT$x[c(4)] = 2.15

G85Rcr = data.frame(Genotype = "G85Rcr",
                   Boutons = na.omit(C$G85Rcr),
                   x = 3)
G85Rcr = G85Rcr[order(G85Rcr$Boutons),]
G85Rcr$x[c(4,10)] = 2.875
G85Rcr$x[c(5,9)] = 3.125
G85Rcr$x[c(1)] = 2.925
G85Rcr$x[c(2)] = 3.075
G85Rcr$x[c(6)] = 2.75
G85Rcr$x[c(7)] = 3.25

A4V = data.frame(Genotype = "A4V",
                   Boutons = na.omit(C$A4V),
                   x = 4)
A4V = A4V[order(A4V$Boutons),]
A4V$x[c(4, 8)] = 3.875
A4V$x[c(5, 9)] = 4.125
A4V$x[c(6)] = 3.75
A4V$x[c(7)] = 4.25
A4V$x[c(1)] = 3.925
A4V$x[c(2)] = 4.075


C = rbind(W1118, rbind(SILENT,
                       rbind(G85Rcr, A4V)))

C$Genotype = factor(C$Genotype, levels = c("W1118", "SILENT", "G85Rcr", "A4V"))

C.labels = c("w1118", "WT(cr)", "G85R(cr)", "A4V")
C.cols = std.colors[C.labels]
     
```


```{r echo = F, fig.width=10, fig.height=7}
 
par(mar = c(5,7,5,2))
      
plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 150),
     xlim = c(.5, 4.5),
     xaxt = "n",
     ylab="Bouton Number",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Pharate muscle VM NMJ bouton counts")
      
boxplot(C$Boutons~C$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 150),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(C.cols, .25),
        add = T, 
        xaxt = "n",
        frame = F)


points(x=as.numeric(C$x),
       y=C$Boutons,
       cex=2.25,
       pch=21,
       bg=C.cols[C$Genotype])
      
axis(side = 1,
     at = c(1:4),
     lwd = 3,
     line = -.9,
     labels = C.labels,
     xpd = T)


axis(side = 2,
     at = c(0, 50, 100, 150),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)

lines(x = c(0.35,4.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

lines(x = c(1, 4),
     y = c(140,140),
     lwd = 3)

lines(x = c(3, 3),
     y = c(60,140),
     lwd = 3)

lines(x = c(1, 1),
     y = c(120,140),
     lwd = 3)

lines(x = c(2, 2),
     y = c(120,140),
     lwd = 3)

lines(x = c(4, 4),
     y = c(120,140),
     lwd = 3)

text(x = 3,
     y = 145,
    labels = "*",
    cex = 2)


```


```{r echo = F, warning=F}

#Kruskal-Wallis test
#Non-parametric → no normality assumption
#Works well for small sample sizes
#Robust to unequal variance
#Standard for this kind of biological data

SummaryStatsC = aggregate(Boutons ~ Genotype, data = C, function(x) {
  c(n = length(x),
    median = median(x),
    mean = mean(x),
    sd = sd(x))
})

Summary =  data.frame(SummaryStatsC$Boutons)

row.names(Summary) = SummaryStatsC$Genotype

Summary

kruskal.test(Boutons ~ Genotype, data = C)

pairwise.wilcox.test(C$Boutons, C$Genotype, p.adjust.method = "BH")

print("There was a significant difference in the number of boutons detected in G85R with every other group tested.")
```


##Larval NMJ 6/7

```{r echo = F}

D = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1D NMJ_quantification_larva.csv")

w1118 = data.frame(Genotype = "w1118",
                   Boutons = na.omit(D$w1118),
                   x = 1)
w1118 = w1118[order(w1118$Boutons),]
w1118$x[c(3, 8, 15)] = .85
w1118$x[c(4, 9, 16)] = 1.15
w1118$x[c(5, 10, 13)] = .7
w1118$x[c(6, 11, 14)] = 1.3


LoxP = data.frame(Genotype = "LoxP",
                   Boutons = na.omit(D$LoxP),
                   x = 2)
LoxP = LoxP[order(LoxP$Boutons),]
LoxP$x[c(3)] = 1.85
LoxP$x[c(4)] = 2.15
LoxP$x[c(5)] = 1.7
LoxP$x[c(6)] = 2.3

G85R_HR = data.frame(Genotype = "G85R_HR",
                   Boutons = na.omit(D$G85R_HR),
                   x = 3)
G85R_HR = G85R_HR[order(G85R_HR$Boutons),]
G85R_HR$x[c(2, 5, 12)] = 2.85
G85R_HR$x[c(3, 6, 13)] = 3.15
G85R_HR$x[c(7, 10)] = 2.7
G85R_HR$x[c(8, 11)] = 3.3

silent = data.frame(Genotype = "silent",
                   Boutons = na.omit(D$silent),
                   x = 4)
silent = silent[order(silent$Boutons),]
silent$x[c(2, 5, 12, 14, 16)] = 3.85
silent$x[c(3, 6, 11, 15, 17)] = 4.15
silent$x[c(7, 10)] = 3.7
silent$x[c(8, 9)] = 4.3

G85R_WG = data.frame(Genotype = "G85R_WG",
                   Boutons = na.omit(D$G85R_WG),
                   x = 5)
G85R_WG = G85R_WG[order(G85R_WG$Boutons),]
G85R_WG$x[c(2)] = 4.85
G85R_WG$x[c(3)] = 5.15
G85R_WG$x[c(4)] = 4.7
G85R_WG$x[c(5)] = 5.3

A4V = data.frame(Genotype = "A4V",
                   Boutons = na.omit(D$A4V),
                   x = 6)
A4V = A4V[order(A4V$Boutons),]
A4V$x[c(1, 3)] = 5.85
A4V$x[c(2, 4)] = 6.15


D = rbind(w1118, rbind(LoxP,
    rbind(G85R_HR, 
    rbind(silent,
    rbind(G85R_WG, A4V)))))

D$Genotype = factor(D$Genotype, levels = c("w1118", "LoxP", "G85R_HR", "silent", "G85R_WG", "A4V"))


D.labels = c("w1118", "WT(LoxP)", "G85R(hr)", "WT(cr)", "G85R(cr)", "A4V")

D.cols = std.colors[D.labels]

```


```{r echo = F, fig.width=10, fig.height=7}
 
par(mar = c(5,7,5,2))
      
plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 150),
     xlim = c(.5, 6.5),
     xaxt = "n",
     ylab="Bouton Number",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Larval muscle 6/7 NMJ bouton counts")
      
boxplot(D$Boutons~D$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 150),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(D.cols, .25),
        add = T, 
        xaxt = "n",
        frame = F)


points(x=as.numeric(D$x),
       y=D$Boutons,
       cex=2.25,
       pch=21,
       bg=D.cols[D$Genotype])
      
axis(side = 1,
     at = c(1:6),
     lwd = 3,
     line = -.9,
     labels = D.labels,
     xpd = T)


axis(side = 2,
     at = c(0, 50, 100, 150),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)

lines(x = c(0.25,6.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)



```


```{r echo =F}

fit <- lm(Boutons ~ Genotype, data = D)

emm <- emmeans(fit, ~ Genotype)

contrast_results = 
  contrast(
    emm,
    method = list(
      "G85R_HR vs LoxP"      = c( 0,-1, 1, 0, 0, 0),
      "G85R_WG vs silent"    = c( 0, 0, 0,-1, 1, 0),
      "A4V vs silent"        = c( 0, 0, 0,-1, 0, 1),
      "A4V vs G85R_WG"       = c( 0, 0, 0, 0,-1, 1)
    )
  )

contrast_results

summary(aov(Boutons ~ Genotype, data = D))

print("Honestly, I think this just doesn't have enough power to detect real , moderate shifts in boutons. The pattern shows a noisy but consistent decrease in boutons for G85R vs control in both hr and cr. It also shows a noisy but strong increase for A4V. However, nothing was significant to a adj. p of .05 when fit to a linear model and performing planned contrasts (description of the test from chat: Pairwise genotype comparisons were performed as planned contrasts from a linear model and evaluated using t-tests with model-derived standard errors.). There was a significant main effect of genotype observed with the anova reflecting a difference in the genotypes, but by itself this isn't exactly useful information by itself. ")



```


##Larval locomotion


```{r echo = F, fig.width=10, fig.height=7}

E = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1E LarvalLocomotion.csv")

E = E[c("w1118", "WTLoxP.WTLoxP", "G85R.G85R.HR.", "WT.WT",
        "G85R.G85R.CR.", "A4V.A4V",
        "A4V.X16", "A4V.G85R.CR.")]

E = E[, c("WTLoxP.WTLoxP", "G85R.G85R.HR.", "WT.WT",
        "G85R.G85R.CR.", "A4V.A4V")]

E.labels = c("WT(LoxP)", 
             "G85R(hr)", "WT(cr)", "G85R(cr)", 
             "A4V")

E.cols = std.colors[E.labels]




n.E = setNames(rep(0, ncol(E)), colnames(E))
mean.E = n.E
sd.E = n.E

i=1
while(i<= ncol(E)){
  mean.E[i] = mean(na.omit(E[,i]))
  sd.E[i] = sd(na.omit(E[,i]))
  n.E[i] = (length(na.omit(E[,i])))
  i=i+1
}

stderr.E = sd.E/(sqrt(n.E))

bar.colors = hex(E.cols, .75)

border.colors = E.cols

tick.x = c(.7 + (c(0:4)*1.2))


#Plots mean larval locomotion and the standard error
par(mar = c(8,4,4,2), lwd = 3)

barplot(height = mean.E,
        ylim = c(0, 80),
        col = border.colors,
        las = 2,
        ylab = "Distance traveled (mm)",
        main = "Larval locomotion",
        border = border.colors,
        xaxt = "n")

axis(side = 1,
     at = tick.x,
     lwd = 3,
     line = 0,
     labels = E.labels,
     xpd = T)

axis(side = 2,
     at = c(0,20,40,60,80),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)
lines(x = c(-0.05,6.2),
     y = c(0,0),
     lwd = 3,
     xpd = T)

i = 1
while(i <= length(stderr.E)){
  lines(x = tick.x[c(i,i)],
        y = c(mean.E[i], (mean.E[i] + stderr.E[i])),
        lwd = 3,
        col = border.colors[i])
  
  lines(x = c((tick.x[i] - .25), (tick.x[i] + .25)),
        y = c((mean.E[i] + stderr.E[i]), (mean.E[i] + stderr.E[i])),
        lwd = 3,
        col = border.colors[i])
  
  i = i +1
}


lines(x = c(.7,1.9),
      y = c(66, 66),
      lwd = 3)

text(x = 1.3,
     y = 69,
    labels = "*",
    cex = 2,
    xpd = t)

```


```{r echo = F, warning=F}

Estats = data.frame(Genotype = c(rep(colnames(E)[1], length(na.omit(E[,1]))),
                                 rep(colnames(E)[2], length(na.omit(E[,2]))),
                                 rep(colnames(E)[3], length(na.omit(E[,3]))),
                                 rep(colnames(E)[4], length(na.omit(E[,4]))),
                                 rep(colnames(E)[5], length(na.omit(E[,5])))),
                    Distance = c(na.omit(E[,1]),
                                 na.omit(E[,2]),
                                 na.omit(E[,3]),
                                 na.omit(E[,4]),
                                 na.omit(E[,5])))

SummaryStatsE = aggregate(Distance ~ Genotype, data = Estats, function(x) {
  c(n = length(x),
    median = median(x),
    mean = mean(x),
    sd = sd(x))
})

Summary =  data.frame(SummaryStatsE$Distance)

row.names(Summary) = SummaryStatsE$Genotype

Summary

kruskal.test(Distance ~ Genotype, data = Estats)
pairwise.wilcox.test(Estats$Distance, Estats$Genotype, p.adjust.method = "BH")

print("Wilcoxon rank sum test shows significant difference in distance traveled between G85R(hr) and WT(LoxP), but not between G85R(cr) and WT(cr) or A4V and WT(cr)")

```



##A4V climbing data
##Climbing plots show the mean at each time point, the individual data points overlayed, and the standard error of the mean as a colored ribbon

```{r include = F, echo = F, include = F}

fig2ab = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 2AB.csv", header = T)
fig2ab$Sample = paste(fig2ab$Genotype, paste(fig2ab$Sex, fig2ab$Day))

fig2ab = na.omit(fig2ab)

fig2ab$Genotype[fig2ab$Genotype == "dSodWT"] = "WT(cr)"

fig2ab = subset(fig2ab, Genotype != "w1118")

df_f <- subset(fig2ab, Sex == "F")
df_m <- subset(fig2ab, Sex == "M")

```


```{r echo = F}

ggplot(df_f, aes(x = Day, y = Rate, color = Genotype)) +

  # points
  geom_point(alpha = 0.4, position = position_jitter(width = 0.2)) +

  # mean line
  stat_summary(fun = mean, geom = "line", linewidth = 1.2) +

  # mean points
  stat_summary(fun = mean, geom = "point", size = 2) +

  # CI ribbon
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(fill = Genotype),
               alpha = 0.2, color = NA) +

  scale_color_manual(values = c("WT(cr)" = "royalblue", "A4V" = "deeppink")) +
  scale_fill_manual(values = c("WT(cr)" = "royalblue", "A4V" = "deeppink")) +
  labs(title = "Female") +

  theme_classic()

```


```{r echo = F}

#2 way anova

model_f <- aov(Rate ~ Genotype * Day, data = df_f)
summary(model_f)


print("There was a significant effect of genotype on climbing velocity and a significant effect of age (day), indicating that genotypes differ in overall performance and that climbing declines over time. However, there was no significant genotype-by-day interaction, suggesting that the rate of decline over time does not differ between genotypes.")

model_WTF <- lm(Rate ~ Day, data = subset(df_f, Genotype == "WT(cr)"))

model_A4VF <- lm(Rate ~ Day, data = subset(df_f, Genotype == "A4V"))

summary(model_WTF)
summary(model_A4VF)

print("Both WT F and A4V F had a significant change in velocity over the time course")

```


```{r echo = F}

ggplot(df_m, aes(x = Day, y = Rate, color = Genotype)) +

  # points
  geom_point(alpha = 0.4, position = position_jitter(width = 0.2)) +

  # mean line
  stat_summary(fun = mean, geom = "line", linewidth = 1.2) +

  # mean points
  stat_summary(fun = mean, geom = "point", size = 2) +

  # CI ribbon
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(fill = Genotype),
               alpha = 0.2, color = NA) +

  scale_color_manual(values = c("WT(cr)" = "royalblue", "A4V" = "deeppink")) +
  scale_fill_manual(values = c("WT(cr)" = "royalblue", "A4V" = "deeppink")) +
  labs(title = "Male") +

  theme_classic()

```


```{r echo = F}

#2 way anova

model_m <- aov(Rate ~ Genotype * Day, data = df_m)
summary(model_m)

print("There was a significant effect of genotype on climbing velocity, a significant effect of age (day), indicating that genotypes differ in overall performance and that climbing declines over time, and a significant genotype-by-day interaction, suggesting that the rate of decline over time differ between genotypes.")

model_WTM <- lm(Rate ~ Day, data = subset(df_m, Genotype == "WT(cr)"))

model_A4VM <- lm(Rate ~ Day, data = subset(df_m, Genotype == "A4V"))

summary(model_WTM)
summary(model_A4VM)

print("WT males did not have a significant change in velocity over the time course but A4V M did p = 9.78e-12")
```


##G85R gbb and SaxA eclosion

```{r echo = F, fig.width=15, fig.height=7.5}

ss.gbbEcl = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Ana_gbbEclosionData.csv")


temp.ssgbb = matrix(rep(0, length(unique(ss.gbbEcl$Genotype))*3), nrow = 3)

colnames(temp.ssgbb) = unique(ss.gbbEcl$Genotype)
row.names(temp.ssgbb) = c("total", "Female", "Male")

i=1
while(i<=length(unique(ss.gbbEcl$Genotype))){
  temp = subset(ss.gbbEcl, Genotype == colnames(temp.ssgbb)[i])
  temp.ssgbb["total", i] = sum(temp$TotalEclosed)
  temp.ssgbb["Female", i] = 100*sum(temp$Female)/
    temp.ssgbb["total", i]
  temp.ssgbb["Male", i] = 100*sum(temp$Male)/
    temp.ssgbb["total", i]
  i=i+1
}

par(mar = c(5,4,4,7))
barplot(height = temp.ssgbb[c("Female", "Male"),],
        cex.names = .8,
        cex.axis = .8,
        legend.text = c("Female", "Male"),
        args.legend =  c(x = 8.1, y = 100),
        ylab = "Percent Eclosed")

text(x = c(.7, 1.9, 3.1, 4.3, 5.5, 6.7), y = 105, labels = c(paste0("n = ", temp.ssgbb["total", ])), xpd = T)

abline(h = 50, lty = 2, lwd =2, col = "red")
abline(h =0, lty =1)



```


```{r echo = F, fig.width=10, fig.height=7}

gbbEcl = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Ana_gbbEclosionData.csv")

gbbEcl.labels = unique(gbbEcl$Genotype)

gbbEcl.means = rep(0, length(gbbEcl.labels))

i = 1
while(i<=length(gbbEcl.labels)){
  gbbEcl.means[i] = mean(gbbEcl$PercentEclosion[gbbEcl$Genotype == gbbEcl.labels[i]])
  i = i+1
}

stderr.gbbEcl = setNames(c(sd(gbbEcl[gbbEcl==gbbEcl.labels[1],"PercentEclosion"])/4,
                           sd(gbbEcl[gbbEcl==gbbEcl.labels[2],"PercentEclosion"])/4,
                           sd(gbbEcl[gbbEcl==gbbEcl.labels[3],"PercentEclosion"])/4,
                           sd(gbbEcl[gbbEcl==gbbEcl.labels[4],"PercentEclosion"])/4,
                           sd(gbbEcl[gbbEcl==gbbEcl.labels[5],"PercentEclosion"])/4,
                           sd(gbbEcl[gbbEcl==gbbEcl.labels[6],"PercentEclosion"])/4),
                         gbbEcl.labels)



par(mar = c(12,4,4,2))

barplot(height = gbbEcl.means,
        ylim = c(0, 120),
        col = "black",
        las = 2,
        ylab = "Eclosion frequency (%)",
        main = "Eclosion from the pupal case",
        xaxt = "n",
        yaxt = "n")
axis(side = 1,
     at = c(.7 + (c(0:5)*1.2)),
     lwd = 3,
     line = 0,
     labels = gbbEcl.labels,
     xpd = T,
     las = 2)
axis(side = 2,
     at = c(0,20,40,60,80,100),
     lwd = 3,
     line = 0,
     labels = T,
     xpd = T,
     las = 2)


tick.x = c(.7 + (c(0:5)*1.2))

i = 1
while(i <= length(stderr.gbbEcl)){
  if(stderr.gbbEcl[i] > 0){
    lines(x = tick.x[c(i,i)],
          y = c(gbbEcl.means[i], (gbbEcl.means[i] + stderr.gbbEcl[i])),
          lwd = 3,
          col = "black")
    
    lines(x = c((tick.x[i] - .25), (tick.x[i] + .25)),
          y = c((gbbEcl.means[i] + stderr.gbbEcl[i]), (gbbEcl.means[i] + stderr.gbbEcl[i])),
          lwd = 3,
          col = "black")  
  }
  i = i +1
}


```


```{r echo = F, include = T, warning=F}

#Why Fisher’s test fits this dataset

#counts (male + female = total eclosed)
#effectively success vs failure (eclosed vs not eclosed — even if “not eclosed” isn’t explicitly listed, you can infer totals)
#small number of replicates (4 broods)
#extreme values (lots of zeros)

#This is exactly where Fisher’s test shines:

#robust to small sample sizes
#handles extreme proportions (0% vs high %)
#no distribution assumptions

##Another Fishers Test


fet.gbbEcl = gbbEcl
fet.gbbEcl$eclosed = fet.gbbEcl$TotalEclosed
fet.gbbEcl$not.eclosed = fet.gbbEcl$TotalPupae - fet.gbbEcl$TotalEclosed

fet.gbbEcl <- aggregate(cbind(eclosed, not.eclosed) ~ Genotype, data = fet.gbbEcl, sum)

# total and proportion
fet.gbbEcl$total <- fet.gbbEcl$eclosed + fet.gbbEcl$not.eclosed
fet.gbbEcl$prop  <- fet.gbbEcl$eclosed / fet.gbbEcl$total

comparisons <- list(
  c("WT(LoxP/cr)", "G85R(hr/cr)"),
  c("WT(LoxP/cr)", "OK371-gbb; WT(LoxP/cr)"),
  c("WT(LoxP/cr)", "OK371-SaxA; WT(LoxP/cr)"),
  c("G85R(hr/cr)", "OK371-gbb; G85R(hr/cr)"),
  c("G85R(hr/cr)", "OK371-SaxA; G85R(hr/cr)")
)

results <- data.frame()


#pair is like i, it starts at 1 and increases until it is larger than the length of comparisons
for (pair in comparisons) {
  
  g1 <- pair[1]
  g2 <- pair[2]
  
  d1 <- fet.gbbEcl[fet.gbbEcl$Genotype == g1, ]
  d2 <- fet.gbbEcl[fet.gbbEcl$Genotype == g2, ]
  
  mat <- matrix(c(
    d1$eclosed, d1$not.eclosed,
    d2$eclosed, d2$not.eclosed
  ), nrow = 2, byrow = TRUE)
  
  ft <- fisher.test(mat)
  
  results <- rbind(results, data.frame(
    Genotype1 = g1,
    n1 = d1$total,
    prop1 = d1$prop,
    Genotype2 = g2,
    n2 = d2$total,
    prop2 = d2$prop,
    diff = d1$prop - d2$prop,
    p.value = ft$p.value
  ))
}

results$padj <- p.adjust(results$p.value, method = "BH")

results$prop1 <- round(results$prop1 * 100, 1)
results$prop2 <- round(results$prop2 * 100, 1)
results$diff  <- round(results$diff * 100, 1)
results$p.value <- signif(results$p.value, 3)
results$padj    <- signif(results$padj, 3)

results.gbbeclosion = results


knitr::kable(results.gbbeclosion,
             caption = 'All fishers exact tests for gbb eclosion data.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(results.gbbeclosion))[results.gbbeclosion$padj <= .05], bold = T, color = "black", background = "gold")

```


##Leg Morphology Section

```{r echo = F, warning=F}

UB = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/UBLegMorphologyData.csv")

score_cols = colnames(UB)[c(7:9,12:14)]
score_cols = colnames(UB)[c(8:14)]
print("used all categores related to the neuronal phenotypes, not those for leg morphology")

UB$score <- rowSums(UB[, score_cols], na.rm = TRUE)


print("wilcoxon rank sum test to see if the overall scores for legs in one group are higher than those of the other group")
wilcox.test(score ~ Genotype, data = UB)

binary_cols <- score_cols  # adjust if needed

feature_results <- data.frame()

for (feature in binary_cols) {
  
  tab <- table(UB$Genotype, UB[[feature]])
  
  ft <- fisher.test(tab)
  
  feature_results <- rbind(feature_results, data.frame(
    Feature = feature,
    p.value = ft$p.value
  ))
}

print("fishers exact test for each individual feature used")
feature_results$padj <- p.adjust(feature_results$p.value, method = "BH")

feature_results
```


```{r echo = F, fig.height=5, fig.width=6, warning=F}

df = UB

library(cluster)
library(vegan)

bin_df <- df[, binary_cols]

# Manhattan (Hamming-like)
dist_matrix <- dist(bin_df, method = "manhattan")


print("this plots the distance matrix for the binary leg data in 2D space")
# Ordination
pcoa <- cmdscale(dist_matrix, k = 2)

par(mar = c(5,4,4,7))

plot(pcoa[,1], pcoa[,2],
     col = ifelse(df$Genotype == "WT", "royalblue", "deeppink"),
     pch = 19,
     main = "Phenotype similarity (binary features)",
     xlab = "Axis 1", ylab = "Axis 2")

legend(x = 3.5,
       y = 2,
       xpd = T,
       legend = unique(df$Genotype),
       col = c("royalblue", "deeppink"),
       pch = 19,
       bty = "n")

print("this test is done to see if the within group variation is smaller than the variation between groups")
# PERMANOVA
adonis2(dist_matrix ~ Genotype, data = df)


print("These tests are done to see if there is a significant difference between the within group variations for WT and G85R")
betadisper(dist_matrix, df$Genotype)
anova(betadisper(dist_matrix, df$Genotype))

```


```{r echo = F, warning=F}

temp = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/UBLegMorphologyData.csv")

temp$Genotype[temp$Genotype == "WT"] = "WT(cr)"
temp$Genotype[temp$Genotype == "G85R"] = "G85R(cr)"

temp$Genotype = factor(temp$Genotype, levels = unique(temp$Genotype))

morph.data = temp[,c(1:2,7:9,12:14)]
morph.data = temp[,c(1:2,8:14)]


sum.morph = setNames(apply(morph.data[,3:ncol(morph.data)], MARGIN = 1, sum), morph.data$Genotype)

#Wilcoxon rank sum test, are the values in one group higher than the other?
wrst = wilcox.test(sum.morph ~ temp$Genotype)
pval = wrst[[3]]

par(mar = c(5,7,5,2))
      
plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 7),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= "Morphology Score",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Leg Axon Morphology")

grid(nx=NA, ny=NULL, col = "grey50", lwd = .75)
      
boxplot(sum.morph~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 6),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        yaxt = "n",
        frame = F)


  axis(side = 1,
       at = c(1,2),
       lwd = 0,
       line = -.45,
       labels = unique(temp$Genotype),
       xpd = T)
  
    lines(x = c(0.4, 2.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)


beeswarm(sum.morph~as.numeric(temp$Genotype),
       pch = 21,
       cex = 1.5,
       spacing = 1.0,
       method = "swarm",
       bg = std.colors[as.character(unique(temp$Genotype))],
       add = T)

      axis(side = 2,
       at = c(0:7),
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)
  
  text(x = 2.25,
       y = 7.5,
       label = paste0("pval = ", signif(pval, 3)))
  
```


```{r echo = F}
library(ggplot2)
library(ggridges)

#best.combo = data_matrix[,c(2,3,4,7,8,9)]


temp = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/UBLegMorphologyData.csv")

temp.titles = c("Is the leg short?",
                "Is there scrunching or curving?",
                "Is the axonal bundle straight?",
                "Does the axonal bundle split or are there chunks missing?",
                "Does the branching look organized?",
                "Is the branching following a pattern tiling?",
                "Is there branching in all aspects of the femur?",
                "Is the mitochonria in the entire axonal bundle?",
                "Do all branches have mitochondria?")


best.titles = temp.titles[c(2,3,4,7,8,9)]

temp$Genotype[temp$Genotype == "WT"] = "WT(cr)"
temp$Genotype[temp$Genotype == "G85R"] = "G85R(cr)"

temp$Genotype = factor(temp$Genotype, levels = unique(temp$Genotype))

morph.data = temp[,c(1:2,7:9,12:14)]

WT.morphdata = morph.data[morph.data$Genotype == "WT(cr)",]
WT.morphdata = t(WT.morphdata[,3:8])
WT.morphdata = as.data.frame(WT.morphdata)
WT.morphdata$Morph = best.titles
wt.morphplot = data.frame(Morph = as.character(),
                          Score = as.numeric())
i=1
while(i<=9){
  temp.morph = data.frame(Morph = WT.morphdata$Morph,
                          Score = WT.morphdata[,i])
  wt.morphplot = rbind(wt.morphplot, temp.morph)
  i = i +1
}
wt.morphplot$Morph = factor(wt.morphplot$Morph, levels = best.titles)
wt.morphplot$Species = "WT(cr)"


GR.morphdata = morph.data[morph.data$Genotype == "G85R(cr)",]
GR.morphdata = t(GR.morphdata[,3:8])
GR.morphdata = as.data.frame(GR.morphdata)
GR.morphdata$Morph = best.titles
GR.morphplot = data.frame(Morph = as.character(),
                          Score = as.numeric())
i=1
while(i<=9){
  temp.morph = data.frame(Morph = GR.morphdata$Morph,
                          Score = GR.morphdata[,i])
  GR.morphplot = rbind(GR.morphplot, temp.morph)
  i = i +1
}
GR.morphplot$Morph = factor(GR.morphplot$Morph, levels = best.titles)
GR.morphplot$Species = "G85R(cr)"

sum.morph = data.frame(Morph = "Total Score",
                       Score = apply(morph.data[,3:8], MARGIN = 1, sum),
                       Species = morph.data$Genotype)
sum.morph$Species = factor(sum.morph$Species, levels = c("WT(cr)", "G85R(cr)"))

all.morphplot = rbind(wt.morphplot, GR.morphplot)
all.morphplot$Species = factor(all.morphplot$Species, levels = c("WT(cr)", "G85R(cr)"))
#all.morphplot = rbind(all.morphplot, sum.morph)


ggplot(
  all.morphplot, 
  aes(x = `Score`, y = `Morph`)
  ) +
  geom_density_ridges_gradient(
    aes(fill = Species), scale = 1
    ) +
  scale_fill_manual(
    values = alpha(c("royalblue", "orange"), .5)
  ) +
  labs(title = 'Morphology Score')

```


```{r echo = F, fig.width=9}

morph.percents = data.frame(WT = as.numeric(),
                            G85R = as.numeric())
i = 3
while(i<= ncol(morph.data)){
  temp.wt = morph.data[morph.data$Genotype == "WT(cr)",i]
  temp.g85r = morph.data[morph.data$Genotype == "G85R(cr)",i]
  morph.percents[i-2, ] = c((sum(temp.wt)/length(temp.wt)) *100,
                            (sum(temp.g85r)/length(temp.g85r)) *100)
  i = i+1
}

row.names(morph.percents) = colnames(morph.data)[3:ncol(morph.data)]

par(mar = c(5,24,5,10))
barplot(as.matrix(t(morph.percents))[2:1,],
        names.arg = best.titles,
        beside = T,
        horiz = T,
        las = 2,
        col = alpha(std.colors[c("WT(cr)", "G85R(cr)")][2:1], 1),
        xlab = "Percent of Samples with Normal Morphology")
legend(x= 100,
       y = 18,
       legend = c("WT(cr)", "G85R(cr)"), 
       fill = alpha(std.colors[c("WT(cr)", "G85R(cr)")], 1),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

```


##VNC pMad Fluorescence

```{r echo = F, fig.show=T}

library(beeswarm)

temp = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/20250809_AllAnaData.csv", row.names = 1)
temp = temp[temp$Boutons != "x",]
temp$Boutons = as.numeric(temp$Boutons)
temp$Genotype[temp$Genotype == "WT-cr/LoxP"] = "WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "G85R"] = "G85R(cr/hr)"
temp$Genotype[temp$Genotype == "G85Rcr/G85Rhr"] = "G85R(cr/hr)"

temp$Genotype[temp$Genotype == "OK371 > gbb [WT]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371 > gbb [G85R]"] = "OK371-gbb; G85R(cr/hr)"
temp$Genotype[temp$Genotype == "OK371>gbb [WT]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [G85R]"] = "OK371-gbb; G85R(cr/hr)"
temp$Genotype[temp$Genotype == "OK371>gbb [WTcr/LoxP]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [WT-cr/LoxP]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [G85Rcr/G85Rhr]"] = "OK371-gbb; G85R(cr/hr)"
temp$Genotype[temp$Genotype == "OK371>gbb [G85cr/G85Rhr]"] = "OK371-gbb; G85R(cr/hr)"

temp$Genotype[temp$Genotype == "OK371 > saxA [WT]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371 > saxA [G85R]"] = "OK371-saxA; G85R(cr/hr)"
temp$Genotype[temp$Genotype == "OK371> saxA [WT-cr/LoxP]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>saxA G85R"] = "OK371-saxA; G85R(cr/hr)"
temp$Genotype[temp$Genotype == "OK371>saxA [WTcr/LoxP]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>saxA [WT-cr/LoxP]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>saxA [G85Rcr/G85Rhr]"] = "OK371-saxA; G85R(cr/hr)"


temp$Genotype = factor(temp$Genotype, levels = unique(temp$Genotype))

VNC = temp[temp$Data == "VNC",]
VNC$CTCF = VNC$CTCF/1000
geno = unique(VNC$Genotype)


```


```{r echo = F, fig.height=6}



par(mar=c(11.25, 4.1, 4.1, 2))


boxplot(VNC$CTCF~VNC$Genotype,
        xaxt = 'none',
        xlab = NA,
        col = hex(std.colors[unique(VNC$Genotype)], .25),
        ylab = "RFU",
        main = "Larval A3 MN nuclei pMad fluorescence",
        ylim = c(0,1800))

beeswarm(VNC$CTCF~as.numeric(VNC$Genotype),
       pch = 21,
       cex = 1.5,
       spacing = 1.5,
       method = "swarm",
       bg = as.character(std.colors[unique(VNC$Genotype)]),
       add = T)


axis(side = 1,
     at = c(1:length(unique(VNC$Genotype))),
     labels = c(unique(VNC$Genotype)),
     cex = .75,
     las = 2,
     lwd = 0,
     line = -.75)

lines(x = c(1,2),
      y = c(1100, 1100),
      lwd = 1.5)

text(x = 1.5,
     y = 1175,
    labels = "*",
    cex = 1.5,
    xpd = t)

lines(x = c(2,6),
      y = c(1600, 1600),
      lwd = 1.5)

text(x = 4,
     y = 1675,
    labels = "*",
    cex = 1.5,
    xpd = t)

```


```{r echo =F, warning=F}

VNC$Rescue = "Control"
VNC$Rescue[grep(pattern = "gbb", VNC$Genotype)] = "gbb"
VNC$Rescue[grep(pattern = "saxA", VNC$Genotype)] = "saxA"
VNC$Model = "WT"
VNC$Model[grep(pattern = "G85R", VNC$Genotype)] = "G85R"

model <- lm(CTCF ~ Rescue * Model + Experiment, data=VNC)
anova(model)

emmeans(model, pairwise ~ Model | Rescue)

emmeans(model, pairwise ~ Rescue | Model)

print("Pairwise comparisons were performed using estimated marginal means derived from a linear model, with Tukey-adjusted p-values.")

```


##A4V saxA and gbb sex specific climbing data

```{r include = F, echo = F}

fig4ab = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 4AB.csv", header = T)


#keep = rep(T, nrow(fig4ab))
#keep[fig4ab$day == 2] = F


# = fig4ab[keep,]

female = fig4ab[grep('f', fig4ab$vial_ID),]
female = female[female$day >2,]

male = fig4ab[grep('m', fig4ab$vial_ID),]

f.con = na.omit(female[female$geno == 'silent' | female$geno == 'A4V',])
f.cha = na.omit(female[grep('cha', female$geno),])

fallcha = rbind(f.con, f.cha)
fallcha$sample = paste0(fallcha$geno, fallcha$day)

fcha.mean = unique(fallcha[,c('geno', 'day')])
fcha.mean$sample = paste0(fcha.mean$geno, fcha.mean$day)
fcha.mean$mean = 0

i=1
while(i<=nrow(fcha.mean)){
  fcha.mean$mean[i] = mean(fallcha[fallcha$sample == fcha.mean$sample[i], 'corrected.values'])
  i = i+1
}

fcha.mean = fcha.mean[order(fcha.mean$day),]


m.con = na.omit(male[male$geno == 'silent' | male$geno == 'A4V',])
m.cha = na.omit(male[grep('cha', male$geno),])

mallcha = rbind(m.con, m.cha)
mallcha$sample = paste0(mallcha$geno, mallcha$day)

mcha.mean = unique(mallcha[,c('geno', 'day')])
mcha.mean$sample = paste0(mcha.mean$geno, mcha.mean$day)
mcha.mean$mean = 0

i=1
while(i<=nrow(mcha.mean)){
  mcha.mean$mean[i] = mean(mallcha[mallcha$sample == mcha.mean$sample[i], 'corrected.values'])
  i = i+1
}

mcha.mean = mcha.mean[order(mcha.mean$day),]



```


```{r echo = F, include = F}

female = fig4ab[grep('f', fig4ab$vial_ID),]
female$sample = paste0(female$geno, female$day)

fmean = unique(female[,c('geno', 'day', 'sample')])
fstd = fmean
eb = fmean
fmean$mean = 0
fstd$std = 0

i=1
while(i<=nrow(fmean)){
  fmean$mean[i] = mean(na.omit(female[female$sample == fmean$sample[i], 'corrected.values']))
  
  fstd$std[i] = (var(na.omit(female[female$sample == fmean$sample[i], 'corrected.values'])))
  
  i = i+1
}

flm = data.frame(geno = unique(female$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(flm)){
  temp=na.omit(female[female$geno == flm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  flm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  flm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(flm) = flm$geno

fmean = fmean[order(fmean$day),]
fstd = fstd[order(fstd$day),]
eb = eb[order(eb$day),]

eb$y0 = fmean$mean - fstd$std
eb$y1 = fmean$mean + fstd$std

```


```{r echo = F, include = F}

male = fig4ab[grep('m', fig4ab$vial_ID),]
male$sample = paste0(male$geno, male$day)

mmean = unique(male[,c('geno', 'day', 'sample')])
mstd = mmean
eb = mmean
mmean$mean = 0
mstd$std = 0

i=1
while(i<=nrow(mmean)){
  mmean$mean[i] = mean(na.omit(male[male$sample == mmean$sample[i], 'corrected.values']))
  
  mstd$std[i] = (var(na.omit(male[male$sample == mmean$sample[i], 'corrected.values'])))
  
  i = i+1
}

mlm = data.frame(geno = unique(male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(mlm)){
  temp=na.omit(male[male$geno == mlm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(mlm) = mlm$geno

mmean = mmean[order(mmean$day),]
mstd = mstd[order(mstd$day),]
eb = eb[order(eb$day),]

eb$y0 = mmean$mean - mstd$std
eb$y1 = mmean$mean + mstd$std

```



```{r echo = F, fig.width=10, fig.height=6}

lcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, max(c(fmean$mean, mmean$mean))),
     xlim = c(1,27),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Females",
     frame.plot = F)

#axis(1, at = fmean[fmean$geno == 'silent', 'day'])
axis(1, at = 1:27, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:13), tick = F, line = -0.75)

axis(2, at = 0:3, label = F, lwd = 3, line = -0.05)
axis(2, at = 0:3, tick = F, las = 1, line = -0.05)

lines(x=c(0,0),
      y=c(0,3.5),
      lwd = 3,
      xpd = T)

lines(x=c(0,27),
      y=c(0,0),
      lwd = 3,
      xpd = T)


lines(x = fmean[fmean$geno == 'silent', 'day'],
      y = fmean[fmean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = fmean[fmean$geno == 'A4V', 'day'],
      y = fmean[fmean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = fmean[fmean$geno == 'silentcha', 'day'],
      y = fmean[fmean$geno == 'silentcha', 'mean'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = fmean[fmean$geno == 'A4Vcha', 'day'],
      y = fmean[fmean$geno == 'A4Vcha', 'mean'],
      type = 'l',
      col = pcol[4],
      lwd = 2)



points(x = fmean[fmean$geno == 'silent', 'day'],
       y = fmean[fmean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4V', 'day'],
       y = fmean[fmean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = fmean[fmean$geno == 'silentcha', 'day'],
       y = fmean[fmean$geno == 'silentcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4Vcha', 'day'],
       y = fmean[fmean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)




points(x = fmean[fmean$geno == 'silent', 'day'],
       y = fmean[fmean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = fmean[fmean$geno == 'A4V', 'day'],
       y = fmean[fmean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = fmean[fmean$geno == 'silentcha', 'day'],
       y = fmean[fmean$geno == 'silentcha', 'mean'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4Vcha', 'day'],
       y = fmean[fmean$geno == 'A4Vcha', 'mean'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)







#legend(x = 29,
#       y = 1.8,
legend("topright",
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V'),
       pch = c(23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5),
                hex(pcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```


```{r echo = F}

library(lme4)
library(lmerTest)


##lmer = linear mixed-effect model
##interaction effect (geno*day) for each genotype, generate a model that estimates velocity at a given time. Treats geno as a factor and time as a number due to the type of data it is (factor and numeric). Change day to factor if you don't want to assume a steady decline or compare rate of decline. That would see if any days are different.

## random effect (1|vial_ID) treats vial_ID as a source of variation that we are not concerned with testing but is potentially adding noise to the model


##the (1 | vial_ID) argument allows the model to treat each vial independently for comparison. For example, if a vial is inherently faster, it doesn't throw off the model for the whole genotype.

##Since day is numeric, the model is testing aging rate instead of determining if genotypes are different at specific days.

fallcha$geno = factor(fallcha$geno, levels = c("silent", "silentcha", "A4V", "A4Vcha"))

fit <- lmer(
  corrected.values ~ geno * day +
    (1 | vial_ID),
  data = fallcha
)

anova(fit)

summary(fit)



emmeans(fit, ~ geno)                    # overall levels
emtrends(fit, ~ geno, var="day")        # aging rates
emmeans(fit, ~ geno | day,
        at=list(day=unique(fallcha$day)))  # specific timepoints

emm <- emmeans(fit, ~ geno)

pairs(emm, adjust = "holm")

trends <- emtrends(fit, ~ geno, var = "day")

pairs(trends, adjust = "holm")

```


```{r echo = F, fig.width=10, fig.height=6}

lcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, max(c(fmean$mean, mmean$mean))),
     xlim = c(6,28),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Females",
     frame.plot = F)

#axis(1, at = fmean[fmean$geno == 'silent', 'day'])
axis(1, at = 1:27, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(3:14), tick = F, line = -0.75)

axis(2, at = 0:3, label = F, lwd = 3, line = -0.05)
axis(2, at = 0:3, tick = F, las = 1, line = -0.05)

lines(x=c(0,0),
      y=c(0,3.5),
      lwd = 3,
      xpd = T)

lines(x=c(5,28),
      y=c(0,0),
      lwd = 3,
      xpd = T)


lines(x = fmean[fmean$geno == 'silent', 'day'],
      y = fmean[fmean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = fmean[fmean$geno == 'A4V', 'day'],
      y = fmean[fmean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)




points(x = fmean[fmean$geno == 'silent', 'day'],
       y = fmean[fmean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4V', 'day'],
       y = fmean[fmean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)





points(x = fmean[fmean$geno == 'silent', 'day'],
       y = fmean[fmean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = fmean[fmean$geno == 'A4V', 'day'],
       y = fmean[fmean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)








#legend(x = 29,
#       y = 1.8,
legend(x = 25, y = 3,
       legend = c('Silent', 'A4V'),
       pch = c(23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```


```{r echo = F, fig.width=10, fig.height=6}

lcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, max(c(fmean$mean, mmean$mean))),
     xlim = c(1,27),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Males",
     frame.plot = F)


#axis(1, at = fmean[fmean$geno == 'silent', 'day'])
axis(1, at = 1:27, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:13), tick = F, line = -0.75)

axis(2, at = 0:3, label = F, lwd = 3, line = -0.05)
axis(2, at = 0:3, tick = F, las = 1, line = -0.05)

lines(x=c(0,0),
      y=c(0,3.5),
      lwd = 3,
      xpd = T)

lines(x=c(0,27),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = mmean[mmean$geno == 'silent', 'day'],
      y = mmean[mmean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4V', 'day'],
      y = mmean[mmean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)


points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)



points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)






#legend(x = 29,
#       y = 1.8,
legend(x =24, y = 3.5,
       legend = c('Silent', 'A4V'),
       pch = c(23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```




```{r echo = F, fig.width=10, fig.height=6}

lcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, max(c(fmean$mean, mmean$mean))),
     xlim = c(1,27),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Males",
     frame.plot = F)

#axis(1, at = fmean[fmean$geno == 'silent', 'day'])
axis(1, at = 1:27, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:13), tick = F, line = -0.75)

axis(2, at = 0:3, label = F, lwd = 3, line = -0.05)
axis(2, at = 0:3, tick = F, las = 1, line = -0.05)

lines(x=c(0,0),
      y=c(0,3.5),
      lwd = 3,
      xpd = T)

lines(x=c(0,27),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = mmean[mmean$geno == 'silent', 'day'],
      y = mmean[mmean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4V', 'day'],
      y = mmean[mmean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = mmean[mmean$geno == 'silentcha', 'day'],
      y = mmean[mmean$geno == 'silentcha', 'mean'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4Vcha', 'day'],
      y = mmean[mmean$geno == 'A4Vcha', 'mean'],
      type = 'l',
      col = pcol[4],
      lwd = 2)



points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentcha', 'day'],
       y = mmean[mmean$geno == 'silentcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vcha', 'day'],
       y = mmean[mmean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)




points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentcha', 'day'],
       y = mmean[mmean$geno == 'silentcha', 'mean'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vcha', 'day'],
       y = mmean[mmean$geno == 'A4Vcha', 'mean'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)







#legend(x = 29,
#       y = 1.8,
legend("topright",
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5), hex(pcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```




```{r echo = F}


mallcha$geno = factor(mallcha$geno, levels = c("silent", "silentcha", "A4V", "A4Vcha"))

fit <- lmer(
  corrected.values ~ geno * day +
    (1 | vial_ID),
  data = mallcha
)

anova(fit)

summary(fit)



emmeans(fit, ~ geno)                    # overall levels
emtrends(fit, ~ geno, var="day")        # aging rates
emmeans(fit, ~ geno | day,
        at=list(day=unique(mallcha$day)))  # specific timepoints

emm <- emmeans(fit, ~ geno)

pairs(emm, adjust = "holm")

trends <- emtrends(fit, ~ geno, var = "day")

pairs(trends, adjust = "holm")

```


```{r echo = F, fig.width=10, fig.height=6}

lcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, max(c(fmean$mean, mmean$mean))),
     xlim = c(1,27),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Males",
     frame.plot = F)

#axis(1, at = fmean[fmean$geno == 'silent', 'day'])
axis(1, at = 1:27, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:13), tick = F, line = -0.75)

axis(2, at = 0:3, label = F, lwd = 3, line = -0.05)
axis(2, at = 0:3, tick = F, las = 1, line = -0.05)

lines(x=c(0,0),
      y=c(0,3.5),
      lwd = 3,
      xpd = T)

lines(x=c(0,27),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = mmean[mmean$geno == 'silent', 'day'],
      y = mmean[mmean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4V', 'day'],
      y = mmean[mmean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = mmean[mmean$geno == 'silentok', 'day'],
      y = mmean[mmean$geno == 'silentok', 'mean'],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4Vok', 'day'],
      y = mmean[mmean$geno == 'A4Vok', 'mean'],
      type = 'l',
      col = pcol[6],
      lwd = 2)



points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentok', 'day'],
       y = mmean[mmean$geno == 'silentok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vcha', 'day'],
       y = mmean[mmean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)




points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentok', 'day'],
       y = mmean[mmean$geno == 'silentok', 'mean'],
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vok', 'day'],
       y = mmean[mmean$geno == 'A4Vok', 'mean'],
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)







#legend(x = 29,
#       y = 1.8,
legend("topright",
       legend = c('Silent', 'A4V',
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[5], .5), hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```



```{r include = F, echo = F}

fig4cd = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 4CD.csv", header = T)

fig4cd = fig4cd[fig4cd$day != 9,]

female = fig4cd[grep('f', fig4cd$vial_ID),]
male = fig4cd[grep('m', fig4cd$vial_ID),]

f.con = na.omit(female[female$geno == 'silent' | female$geno == 'A4V',])
f.cha = na.omit(female[grep('cha', female$geno),])

fallcha = rbind(f.con, f.cha)
fallcha$sample = paste0(fallcha$geno, fallcha$day)

fcha.mean = unique(fallcha[,c('geno', 'day')])
fcha.mean$sample = paste0(fcha.mean$geno, fcha.mean$day)
fcha.mean$mean = 0

i=1
while(i<=nrow(fcha.mean)){
  fcha.mean$mean[i] = mean(fallcha[fallcha$sample == fcha.mean$sample[i], 'corrected.values'])
  i = i+1
}

fcha.mean = fcha.mean[order(fcha.mean$day),]


m.con = na.omit(male[male$geno == 'silent' | male$geno == 'A4V',])
m.cha = na.omit(male[grep('cha', male$geno),])

mallcha = rbind(m.con, m.cha)
mallcha$sample = paste0(mallcha$geno, mallcha$day)

mcha.mean = unique(mallcha[,c('geno', 'day')])
mcha.mean$sample = paste0(mcha.mean$geno, mcha.mean$day)
mcha.mean$mean = 0

i=1
while(i<=nrow(mcha.mean)){
  mcha.mean$mean[i] = mean(mallcha[mallcha$sample == mcha.mean$sample[i], 'corrected.values'])
  i = i+1
}

mcha.mean = mcha.mean[order(mcha.mean$day),]



```


```{r echo = F, include = F}

female = fig4cd[grep('f', fig4cd$vial_ID),]
female$sample = paste0(female$geno, female$day)

fmean = unique(female[,c('geno', 'day', 'sample')])
fstd = fmean
eb = fmean
fmean$mean = 0
fstd$std = 0

i=1
while(i<=nrow(fmean)){
  fmean$mean[i] = mean(na.omit(female[female$sample == fmean$sample[i], 'corrected.values']))
  
  fstd$std[i] = (var(na.omit(female[female$sample == fmean$sample[i], 'corrected.values'])))
  
  i = i+1
}

flm = data.frame(geno = unique(female$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(flm)){
  temp=na.omit(female[female$geno == flm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  flm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  flm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(flm) = flm$geno

fmean = fmean[order(fmean$day),]
fstd = fstd[order(fstd$day),]
eb = eb[order(eb$day),]

eb$y0 = fmean$mean - fstd$std
eb$y1 = fmean$mean + fstd$std

```


```{r echo = F, include = F}

male = fig4cd[grep('m', fig4cd$vial_ID),]
male$sample = paste0(male$geno, male$day)

mmean = unique(male[,c('geno', 'day', 'sample')])
mstd = mmean
eb = mmean
mmean$mean = 0
mstd$std = 0

i=1
while(i<=nrow(mmean)){
  mmean$mean[i] = mean(na.omit(male[male$sample == mmean$sample[i], 'corrected.values']))
  
  mstd$std[i] = (var(na.omit(male[male$sample == mmean$sample[i], 'corrected.values'])))
  
  i = i+1
}

mlm = data.frame(geno = unique(male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(mlm)){
  temp=na.omit(male[male$geno == mlm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(mlm) = mlm$geno

mmean = mmean[order(mmean$day),]
mstd = mstd[order(mstd$day),]
eb = eb[order(eb$day),]

eb$y0 = mmean$mean - mstd$std
eb$y1 = mmean$mean + mstd$std

```



```{r echo = F, fig.width=10, fig.height=6}

lcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, max(c(fmean$mean, mmean$mean))),
     xlim = c(1,32),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Females",
     frame.plot = F)

#axis(1, at = fmean[fmean$geno == 'silent', 'day'])
axis(1, at = 1:32, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:16), tick = F, line = -0.75)

axis(2, at = 0:2, label = F, lwd = 3, line = -0.2)
axis(2, at = 0:2, tick = F, las = 1, line = -0.2)

lines(x=c(0,0),
      y=c(0,2),
      lwd = 3,
      xpd = T)

lines(x=c(0,32),
      y=c(0,0),
      lwd = 3,
      xpd = T)


lines(x = fmean[fmean$geno == 'silent', 'day'],
      y = fmean[fmean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = fmean[fmean$geno == 'A4V', 'day'],
      y = fmean[fmean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = fmean[fmean$geno == 'silentcha', 'day'],
      y = fmean[fmean$geno == 'silentcha', 'mean'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = fmean[fmean$geno == 'A4Vcha', 'day'],
      y = fmean[fmean$geno == 'A4Vcha', 'mean'],
      type = 'l',
      col = pcol[4],
      lwd = 2)



points(x = fmean[fmean$geno == 'silent', 'day'],
       y = fmean[fmean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4V', 'day'],
       y = fmean[fmean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = fmean[fmean$geno == 'silentcha', 'day'],
       y = fmean[fmean$geno == 'silentcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4Vcha', 'day'],
       y = fmean[fmean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)




points(x = fmean[fmean$geno == 'silent', 'day'],
       y = fmean[fmean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = fmean[fmean$geno == 'A4V', 'day'],
       y = fmean[fmean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = fmean[fmean$geno == 'silentcha', 'day'],
       y = fmean[fmean$geno == 'silentcha', 'mean'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4Vcha', 'day'],
       y = fmean[fmean$geno == 'A4Vcha', 'mean'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)







#legend(x = 29,
#       y = 1.8,
legend("topright",
       legend = c('Silent', 'A4V',
                  'ChAT>saxA;Silent', 'ChAT>saxA;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5), hex(pcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```


```{r echo = F, fig.width=10, fig.height=6}

lcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, max(c(fmean$mean, mmean$mean))),
     xlim = c(1,32),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Females",
     frame.plot = F)

#axis(1, at = fmean[fmean$geno == 'silent', 'day'])
axis(1, at = 1:32, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:16), tick = F, line = -0.75)

axis(2, at = 0:2, label = F, lwd = 3, line = -0.2)
axis(2, at = 0:2, tick = F, las = 1, line = -0.2)

lines(x=c(0,0),
      y=c(0,2),
      lwd = 3,
      xpd = T)

lines(x=c(0,32),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = fmean[fmean$geno == 'silent', 'day'],
      y = fmean[fmean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = fmean[fmean$geno == 'A4V', 'day'],
      y = fmean[fmean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)



lines(x = fmean[fmean$geno == 'silentok', 'day'],
      y = fmean[fmean$geno == 'silentok', 'mean'],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = fmean[fmean$geno == 'A4Vok', 'day'],
      y = fmean[fmean$geno == 'A4Vok', 'mean'],
      type = 'l',
      col = pcol[6],
      lwd = 2)



points(x = fmean[fmean$geno == 'silent', 'day'],
       y = fmean[fmean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4V', 'day'],
       y = fmean[fmean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = fmean[fmean$geno == 'silentok', 'day'],
       y = fmean[fmean$geno == 'silentok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4Vok', 'day'],
       y = fmean[fmean$geno == 'A4Vok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)



points(x = fmean[fmean$geno == 'silent', 'day'],
       y = fmean[fmean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = fmean[fmean$geno == 'A4V', 'day'],
       y = fmean[fmean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)

points(x = fmean[fmean$geno == 'silentok', 'day'],
       y = fmean[fmean$geno == 'silentok', 'mean'],
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4Vok', 'day'],
       y = fmean[fmean$geno == 'A4Vok', 'mean'],
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)






#legend(x = 29,
#       y = 1.8,
legend("topright",
       legend = c('Silent', 'A4V',
                  'OK371>saxA;Silent', 'OK371>saxA;A4V'),
       pch = c(23,21,23,21,23,21, 26, 32),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[5], .5), hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```





```{r echo = F, fig.width=10, fig.height=6}

lcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, max(c(fmean$mean, mmean$mean))),
     xlim = c(1,32),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Males",
     frame.plot = F)

#axis(1, at = fmean[fmean$geno == 'silent', 'day'])
axis(1, at = 1:32, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:16), tick = F, line = -0.75)

axis(2, at = 0:2, label = F, lwd = 3, line = -0.2)
axis(2, at = 0:2, tick = F, las = 1, line = -0.2)

lines(x=c(0,0),
      y=c(0,2),
      lwd = 3,
      xpd = T)

lines(x=c(0,32),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = mmean[mmean$geno == 'silent', 'day'],
      y = mmean[mmean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4V', 'day'],
      y = mmean[mmean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = mmean[mmean$geno == 'silentcha', 'day'],
      y = mmean[mmean$geno == 'silentcha', 'mean'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4Vcha', 'day'],
      y = mmean[mmean$geno == 'A4Vcha', 'mean'],
      type = 'l',
      col = pcol[4],
      lwd = 2)



points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentcha', 'day'],
       y = mmean[mmean$geno == 'silentcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vcha', 'day'],
       y = mmean[mmean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)




points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentcha', 'day'],
       y = mmean[mmean$geno == 'silentcha', 'mean'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vcha', 'day'],
       y = mmean[mmean$geno == 'A4Vcha', 'mean'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)







#legend(x = 29,
#       y = 1.8,
legend("topright",
       legend = c('Silent', 'A4V',
                  'ChAT>saxA;Silent', 'ChAT>saxA;A4V'),
       pch = c(23,21,23,21,23,21, 26, 32),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5),
                hex(pcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```


```{r echo = F, fig, fig.width=10, fig.height=6}

lcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, max(c(fmean$mean, mmean$mean))),
     xlim = c(1,32),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Males",
     frame.plot = F)

#axis(1, at = fmean[fmean$geno == 'silent', 'day'])
axis(1, at = 1:32, label = F, lwd = 3, line = -0.75)
axis(1, at = 2*(1:16), tick = F, line = -0.75)

axis(2, at = 0:2, label = F, lwd = 3, line = -0.2)
axis(2, at = 0:2, tick = F, las = 1, line = -0.2)

lines(x=c(0,0),
      y=c(0,2),
      lwd = 3,
      xpd = T)

lines(x=c(0,32),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = mmean[mmean$geno == 'silent', 'day'],
      y = mmean[mmean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4V', 'day'],
      y = mmean[mmean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = mmean[mmean$geno == 'silentok', 'day'],
      y = mmean[mmean$geno == 'silentok', 'mean'],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4Vok', 'day'],
      y = mmean[mmean$geno == 'A4Vok', 'mean'],
      type = 'l',
      col = pcol[6],
      lwd = 2)



points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentok', 'day'],
       y = mmean[mmean$geno == 'silentok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vcha', 'day'],
       y = mmean[mmean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)




points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentok', 'day'],
       y = mmean[mmean$geno == 'silentok', 'mean'],
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vok', 'day'],
       y = mmean[mmean$geno == 'A4Vok', 'mean'],
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)







#legend(x = 29,
#       y = 1.8,
legend("topright",
       legend = c('Silent', 'A4V',
                  'OK371>saxA;Silent', 'OK371>saxA;A4V'),
       pch = c(23,21,23,21,23,21, 26, 32),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[5], .5),
                hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```


```{r echo = F, fig.width=10, fig.height=6}

fig1f = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1F left_females.csv", header = T)


lcol = c('navy', 'deepskyblue', 'deeppink2', 'grey70')#c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 100),
     xlim = c(1,80),
     ylab = 'Alive Flies (%)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Females",
     frame.plot = F)

#axis(1, at = fmean[fmean$geno == 'silent', 'day'])
axis(1, at = 20*(0:4), label = F, lwd = 3, line = -0.75)
axis(1, at = 20*(0:4), tick = F, line = -0.75)

axis(2, at = 20*(0:5), label = F, lwd = 3, line = -0.8)
axis(2, at = 20*(0:5), tick = F, las = 1, line = -0.8)

lines(x=c(0,0),
      y=c(0,100),
      lwd = 3,
      xpd = T)

lines(x=c(0,85),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = fig1f$day,
      y = fig1f$w1118,
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = fig1f$day,
      y = fig1f$dsodWT,
      type = 'l',
      col = lcol[2],
      lwd = 2)

lines(x = fig1f$day,
      y = fig1f$A4V,
      type = 'l',
      col = lcol[3],
      lwd = 2)

lines(x = fig1f$day,
      y = fig1f$A4V.X16,
      type = 'l',
      col = lcol[4],
      lwd = 2)




#legend(x = 29,
#       y = 1.8,
legend("topright",
       legend = c('w1118', 'WT(cr)',
                  'A4V', 'A4V/X16'),
       pch = 21,
       pt.bg = c(hex(lcol[1], .5), hex(lcol[2], .5), hex(lcol[3], .5),
                hex(lcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```


```{r echo = F, fig.width=10, fig.height=6}

fig1g = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1G left_males.csv", header = T)


lcol = c('navy', 'deepskyblue', 'deeppink2', 'grey70')#c('deepskyblue',  'deeppink', 'darkgreen', "gold3", 'navy', "firebrick")#c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 100),
     xlim = c(1,80),
     ylab = 'Alive Flies (%)',
     xlab = 'Day',
     xaxt = 'n',
     yaxt = 'n',
     main = "Males",
     frame.plot = F)

#axis(1, at = fmean[fmean$geno == 'silent', 'day'])
axis(1, at = 20*(0:4), label = F, lwd = 3, line = -0.75)
axis(1, at = 20*(0:4), tick = F, line = -0.75)

axis(2, at = 20*(0:5), label = F, lwd = 3, line = -0.8)
axis(2, at = 20*(0:5), tick = F, las = 1, line = -0.8)

lines(x=c(0,0),
      y=c(0,100),
      lwd = 3,
      xpd = T)

lines(x=c(0,85),
      y=c(0,0),
      lwd = 3,
      xpd = T)

lines(x = fig1g$day,
      y = fig1g$w1118,
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = fig1g$day,
      y = fig1g$dsodWT,
      type = 'l',
      col = lcol[2],
      lwd = 2)

lines(x = fig1g$day,
      y = fig1g$A4V,
      type = 'l',
      col = lcol[3],
      lwd = 2)

lines(x = fig1g$day,
      y = fig1g$A4V.X16,
      type = 'l',
      col = lcol[4],
      lwd = 2)




#legend(x = 29,
#       y = 1.8,
legend("topright",
       legend = c('w1118', 'WT(cr)',
                  'A4V', 'A4V/X16'),
       pch = 21,
       pt.bg = c(hex(lcol[1], .5), hex(lcol[2], .5), hex(lcol[3], .5),
                hex(lcol[4], .5)),
       bty = 'n',
       y.intersp = 1.2,
       pt.cex = 2,
       xpd=TRUE)

```

##G85R Bouton Counts

```{r echo = F, fig.width=8, fig.height=6}
##Barron's Boutons

bb = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/20260603_BarronBoutonCounts.csv")

bb$Genotype[bb$Genotype == "WT"] = "WTcr/LoxP"
bb$Genotype[bb$Genotype == "G85R"] = "G85Rcr/hr"

bb$Genotype[bb$Genotype == "OK371>gbb WT"] = "OK371>gbb, WT"
bb$Genotype[bb$Genotype == "OK371>gbb G85R"] = "OK371>gbb, G85R"

bb$Genotype[bb$Genotype == "OK371>SaxA WT"] = "OK371>SaxA, WT"
bb$Genotype[bb$Genotype == "OK371>SaxA G85R"] = "OK371>SaxA, G85R"


bb4 = bb[bb$Segment == "A4", c('Genotype', "Boutons")]

bbplot = bb4

bbplot$color = "WT(cr)"
bbplot$color[grep("G85R", bbplot$Genotype)] = "G85R(cr)"
bbplot$color = std.colors[bbplot$color]

bbplot$Genotype = factor(bbplot$Genotype, levels = c("WTcr/LoxP", "G85Rcr/hr",
                                                     "OK371>gbb, WT",
                                                     "OK371>gbb, G85R",
                                                     "OK371>SaxA, WT",
                                                     "OK371>SaxA, G85R"))

par(mar=c(5.1, 4.1, 4.1, 6.3))

boxplot(bbplot$Boutons~bbplot$Genotype,
        at = c(1,2, 4,5, 7,8),
        xaxt = 'none',
        las = 2,
        xlab = "Genotype",
        col = hex(unique(bbplot$color), .25),
        ylab = 'Total Boutons',
        main = "Segment 4 Boutons",
        ylim = c(0,60))


beeswarm(bbplot$Boutons~bbplot$Genotype,
         at = c(1,2, 4,5, 7,8),
       pch = 21,
       cex = 1.0,
       spacing = 1.0,
       method = "swarm",
       bg = unique(bbplot$color),
       add = T)


axis(side = 1,
     at = c(1.5, 4.5, 7.5),
     labels = c("Control", "OK371>gbb", "OK371>SaxA"))

legend('topright',
       inset=c(-0.225,0),
       legend = c('WT(cr/hr)', 'G85R(cr/hr)'), 
       fill = unique(bbplot$color),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

lines(x=c(2,8),
      y=c(60,60))

text(x = 5,
     y = 61,
     labels = "*")

lines(x=c(5,8),
      y=c(57,57))

text(x = 6.5,
     y = 58,
     labels = "*")



```


```{r echo = F, fig.width=8, fig.height=6}

bbplot = bb4

bbplot$color = "grey"
bbplot$color[grep("SaxA", bbplot$Genotype)] = "firebrick"
bbplot$color[grep("gbb", bbplot$Genotype)] = "steelblue"

bbplot$Genotype = factor(bbplot$Genotype, levels = c("WTcr/LoxP",
                                                     "OK371>gbb, WT",
                                                     "OK371>SaxA, WT",
                                                     "G85Rcr/hr",
                                                     "OK371>gbb, G85R",
                                                     "OK371>SaxA, G85R"))

par(mar=c(5.1, 4.1, 4.1, 6.3))

boxplot(bbplot$Boutons~bbplot$Genotype,
        at = c(1,2,3, 5,6,7),
        xaxt = 'none',
        las = 2,
        xlab = "",
        col = hex(c("grey", "steelblue", "firebrick"), .25),
        ylab = 'Total Boutons',
        main = "Segment 4 Boutons",
        ylim = c(0,60))


beeswarm(bbplot$Boutons~bbplot$Genotype,
         at = c(1,2,3, 5,6,7),
       pch = 21,
       cex = 1.0,
       spacing = 1.0,
       method = "swarm",
       bg = c("grey", "steelblue", "firebrick"),
       add = T)


axis(side = 1,
     at = c(2, 6),
     labels = c("WT(cr/hr)", "G85R(cr/hr)"))

legend('topright',
       inset=c(-0.225,0),
       legend = c('Control', 'gbb', "SaxA"), 
       fill = c("grey", "steelblue", "firebrick"),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

lines(x=c(5,7),
      y=c(60,60))

lines(x=c(6,7),
      y=c(57,57))

text(x = 6,
     y = 61,
     labels = "*")

text(x = 6.5,
     y = 58,
     labels = "*")

abline(v = 4, lty = 2)

```


```{r echo = F, warning=F}

print("Generated a linear model using all of the data collected on segment A4 to detect main effects of the Rescue and Model")
A4 <- subset(bb, Segment=="A4")

fit <- lm(Boutons ~ Model * Rescue, data=A4)

anova(fit)

print("The model did not have a significant main effect alone but the rescue did and the interaction effect of the model and rescue was also significant. This indicates that gbb and SaxA are significantly altering bouton number and this effect is different between the genotypes.")

emmeans(fit, pairwise ~ Model * Rescue,
        adjust="tukey")

print("Individual comparisons found that the effects of SaxA in G85R significantly altered bouton counts from G85R in the control and gbb backgrounds")

```


```{r echo = F, fig.width=8, fig.height=6}
##Barron's Boutons

bb5 = bb[bb$Segment == "A5", c('Genotype', "Boutons")]

bbplot = bb5

bbplot$color = "WT(cr)"
bbplot$color[grep("G85R", bbplot$Genotype)] = "G85R(cr)"
bbplot$color = std.colors[bbplot$color]

bbplot$Genotype = factor(bbplot$Genotype, levels = c("WTcr/LoxP", "G85Rcr/hr",
                                                     "OK371>gbb, WT",
                                                     "OK371>gbb, G85R",
                                                     "OK371>SaxA, WT",
                                                     "OK371>SaxA, G85R"))

par(mar=c(5.1, 4.1, 4.1, 6.3))

boxplot(bbplot$Boutons~bbplot$Genotype,
        at = c(1,2, 4,5, 7,8),
        xaxt = 'none',
        las = 2,
        xlab = "Genotype",
        col = hex(unique(bbplot$color), .25),
        ylab = 'Total Boutons',
        main = "Segment 5 Boutons",
        ylim = c(0,60))


beeswarm(bbplot$Boutons~bbplot$Genotype,
         at = c(1,2, 4,5, 7,8),
       pch = 21,
       cex = 1.0,
       spacing = 1.0,
       method = "swarm",
       bg = unique(bbplot$color),
       add = T)


axis(side = 1,
     at = c(1.5, 4.5, 7.5),
     labels = c("Control", "OK371>gbb", "OK371>SaxA"))

legend('topright',
       inset=c(-0.225,0),
       legend = c('WT(cr/hr)', 'G85R(cr/hr)'), 
       fill = unique(bbplot$color),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

lines(x=c(1,2),
      y=c(60,60))

text(x = 1.5,
     y = 61,
     labels = "*")

lines(x=c(1,4),
      y=c(57,57))

text(x = 2.5,
     y = 58,
     labels = "*")

lines(x=c(2,8),
      y=c(54,54))

text(x = 5,
     y = 55,
     labels = "*")



```


```{r echo = F, fig.width=8, fig.height=6}

bbplot = bb5

bbplot$color = "grey"
bbplot$color[grep("SaxA", bbplot$Genotype)] = "firebrick"
bbplot$color[grep("gbb", bbplot$Genotype)] = "steelblue"

bbplot$Genotype = factor(bbplot$Genotype, levels = c("WTcr/LoxP",
                                                     "OK371>gbb, WT",
                                                     "OK371>SaxA, WT",
                                                     "G85Rcr/hr",
                                                     "OK371>gbb, G85R",
                                                     "OK371>SaxA, G85R"))

par(mar=c(5.1, 4.1, 4.1, 6.3))

boxplot(bbplot$Boutons~bbplot$Genotype,
        at = c(1,2,3, 5,6,7),
        xaxt = 'none',
        las = 2,
        xlab = "",
        col = hex(c("grey", "steelblue", "firebrick"), .25),
        ylab = 'Total Boutons',
        main = "Segment 5 Boutons",
        ylim = c(0,60))


beeswarm(bbplot$Boutons~bbplot$Genotype,
         at = c(1,2,3, 5,6,7),
       pch = 21,
       cex = 1.0,
       spacing = 1.0,
       method = "swarm",
       bg = c("grey", "steelblue", "firebrick"),
       add = T)


axis(side = 1,
     at = c(2, 6),
     labels = c("WT(cr/hr)", "G85R(cr/hr)"))

legend('topright',
       inset=c(-0.225,0),
       legend = c('Control', 'gbb', "SaxA"), 
       fill = c("grey", "steelblue", "firebrick"),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

lines(x=c(1,5),
      y=c(60,60))

lines(x=c(1,2),
      y=c(57,57))

lines(x=c(5,7),
      y=c(57,57))

text(x = 2.5,
     y = 61,
     labels = "*")

text(x = 1.5,
     y = 58,
     labels = "*")

text(x = 6,
     y = 58,
     labels = "*")

abline(v = 4, lty = 2)

```


```{r echo = F, warning=F}

print("Generated a linear model using all of the data collected on segment A4 to detect main effects of the Rescue and Model")
A5 <- subset(bb, Segment=="A5")

fit <- lm(Boutons ~ Model * Rescue, data=A5)

anova(fit)

print("Neither the model or rescue had a significant main effect alone but the interaction effect of the model and rescue together was significant. This indicates that gbb and SaxA are significantly altering bouton number differently between the genotypes.")

emmeans(fit, pairwise ~ Model * Rescue,
        adjust="tukey")

print("Individual comparisons found that only WT and G85R have significantly different numbers of boutons at an adj. p of .05. Using a threshold of .1, the comparisons G85R ~ G85R-SaxA and WT ~ WT-gbb were found to have  significantly altered bouton counts")

```


##Aleah embryo data

```{r echo = F, fig.width=5, fig.height=5}

##Embryo survival data
embryo = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/AleahEmbryoSurvivalData.csv")

embryo$Percent = 100*(embryo$Dead/embryo$Total)
plot.embryo = data.frame(WT = embryo$Percent[embryo$Genotype == "Wtcr/loxP"],
                         G85R = embryo$Percent[embryo$Genotype == "G85Rcr/hr"])


embryo.labels = c("WTcr/loxP", "G85Rcr/hr")

embryo.cols = std.colors[c("WT(cr)","G85R(cr)")]

n.embryo = setNames(rep(0, ncol(plot.embryo)), colnames(plot.embryo))
mean.embryo = n.embryo
sd.embryo = n.embryo

i=1
while(i<= ncol(plot.embryo)){
  mean.embryo[i] = mean(na.omit(plot.embryo[,i]))
  sd.embryo[i] = sd(na.omit(plot.embryo[,i]))
  n.embryo[i] = (length(na.omit(plot.embryo[,i])))
  i=i+1
}

stderr.embryo = sd.embryo/(sqrt(n.embryo))

bar.colors = hex(embryo.cols, .75)

border.colors = embryo.cols

tick.x = c(.7 + (c(0:1)*1.2))

par(mar = c(8,4,4,2), lwd = 3)

barplot(height = mean.embryo,
        ylim = c(0, 20),
        col = border.colors,
        las = 2,
        ylab = "% Dead",
        main = "Embryo Survival",
        border = "black",
        xaxt = "n")

axis(side = 1,
     at = tick.x,
     lwd = 3,
     line = 0,
     labels = embryo.labels,
     xpd = T)

axis(side = 2,
     at = c(0,5,10,15,20),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)
lines(x = c(0.1, 2.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

i = 1
while(i <= length(stderr.embryo)){
  lines(x = tick.x[c(i,i)],
        y = c(mean.embryo[i], (mean.embryo[i] + stderr.embryo[i])),
        lwd = 3,
        col = "black")
  
  lines(x = c((tick.x[i] - .25), (tick.x[i] + .25)),
        y = c((mean.embryo[i] + stderr.embryo[i]), (mean.embryo[i] + stderr.embryo[i])),
        lwd = 3,
        col = "black")
  
  i = i +1
}

lines(x = tick.x[1:2],
      y =  c( 15,15),
      lwd = 3)

lines(x = c(1.9,3.1),
      y = c(63, 63),
      lwd = 3)

text(x = 1.3,
     y = 17,
    labels = "*",
    cex = 2,
    xpd = t)





##Stats advice from chatgpt:
##A classic Student’s t-test assumes:
##  Normal distribution (hard to justify with n=4)
##  Similar variances
##  No strong outliers

##Your data violates at least #1 (and maybe #2). With such a small sample, the t-test becomes very unstable — a single value (like that 0) can heavily influence the result.

##Wilcoxon rank-sum test (recommended)
##  Non-parametric → no normality assumption
##  Works well with small n
##  Tests whether one group tends to be larger than the other

##This uses the exact distribution, which is more appropriate than approximations at small n.

wilcox.test(plot.embryo$WT, 
            plot.embryo$G85R, 
            exact = TRUE)

```








